### 2023-11-16 experiment/stimuli_selection/selection.qmd
reduce to 54 stims, adjust timings

---
title: "Stimuli Selection"
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
#| code-fold: true
library(tidyverse)
library(easystats)
library(ggrepel)

naps <- readxl::read_excel("norms_naps.xls",  skip = 3,
                           col_names = c("ID", "Category", "Nr", "Orientation", "Description",
                                         "Women_Valence", "Women_Valence_SD", "Women_AvAp",
                                         "Women_AvAp_SD", "Women_Arousal", "Women_Arousal_SD",
                                         "Men_Valence", "Men_Valence_SD", "Men_AvAp", "Men_AvAp_SD",
                                         "Men_Arousal", "Men_Arousal_SD", "All_Valence",
                                         "All_Valence_SD", "All_AvAp", "All_AvAp_SD",
                                         "All_Arousal", "All_Arousal_SD", "Width",
                                         "Height", "Luminance", "Contrast",
                                         "JPEG_size80", "LABL", "LABA", "LABB", "Entropy",
                                         "Acronym", "X", "X2")) |>
  select(-X, -X2, -Acronym, -ends_with("_SD")) |>
  mutate(ID = paste0(ID, ".jpg"))

napsero <- readxl::read_excel("norms_napsero.XLSX", sheet = "Ratings") |>
  full_join(readxl::read_excel("norms_napsero.XLSX", sheet = "Image parameters"), by = join_by("ID", "Category")) |>
  select(-contains("_SD"), -contains("HoF"), -contains("HoM")) |>
  rename(Orientation = `V/H`,
         Men_Valence = val_M_HeM,
         Women_Valence = val_M_HeF,
         Men_Arousal = aro_M_HeM,
         Women_Arousal = aro_M_HeF) |>
  mutate(across(c("Luminance", "Contrast", "JPEG_size80", "LABL", "LABA", "LABB", "Entropy"), as.numeric),
         Description = NA)

# Join
naps <- filter(naps, ID %in% filter(napsero, Category == "Non-erotic")$ID)
naps[c("Men_Arousal", "Men_Valence", "Women_Valence", "Women_Arousal", "Nr", "Orientation")] <- NA
napsero[napsero$Category== "Non-erotic", "Category"] <- NA

ids <- napsero[is.na(napsero$Category), ]$ID
for(c in c("Category", "Height", "Width", "Luminance", "Contrast",
           "JPEG_size80", "LABL", "LABA", "LABB", "Entropy", "Description")) {
  napsero[napsero$ID == ids, c] <- naps[naps$ID == ids, c]
}

```

## Selection

### Remove couples

```{r}
df <- napsero |>
  filter(Category %in% c("Male", "Female", "Faces", "People")) |>
  mutate(Type = ifelse(Category %in% c("Female", "Male"), "Erotic", "Non-erotic"))
```

### Select based on norms
```{r}
men <- df |>
  filter(Type == "Erotic") |>
  group_by(Category) |>
  slice_max(Men_Arousal, n=6, with_ties = FALSE) |>
  pull(ID) |> 
  c(
    df |>
      filter(Type == "Erotic") |>
      group_by(Category) |>
      slice_min(Men_Arousal, n=2, with_ties = FALSE) |>
      pull(ID))

women <- df |>
  filter(Type == "Erotic", !ID %in% men) |>
  group_by(Category) |>
  slice_max(Women_Arousal, n=6, with_ties = FALSE) |>
  pull(ID) |> 
  c(
    df |>
      filter(Type == "Erotic", !ID %in% men) |>
      group_by(Category) |>
      slice_min(Women_Arousal, n=2, with_ties = FALSE) |>
      pull(ID)
  )

neutral <- df |>
  filter(Type == "Non-erotic", !ID %in% c(men, women)) |>
  mutate(Arousal = (Men_Arousal + Women_Arousal) / 2) |>
  slice_min(Arousal, n=12, with_ties = FALSE) |>
  pull(ID)

pos_arousing <- df |>
  filter(Type == "Non-erotic", !ID %in% c(men, women, neutral)) |>
  mutate(Arousal = (Men_Arousal + Women_Arousal) / 2,
         Valence = (Men_Valence + Women_Valence) / 2) |>
  filter(Valence > 5) |>
  slice_max(Arousal, n=12, with_ties = FALSE) |>
  pull(ID)

selected <- unique(c(men, women, neutral, pos_arousing))

cat(
  paste0("N (men) = ", length(men), "\nN (women) = ", length(women),
       "\nN (neutral) = ", length(neutral),  "\nN (arousing-positive) = ", length(pos_arousing),
       "\nTotal = ", length(selected))
)
```

```{r}
selected
```

### Visualization

```{r}
#| out.width: 1\\textwidth
#| fig.width: 8
#| fig.height: 6
napsero |>
  mutate(Selected = ifelse(ID %in% selected, TRUE, FALSE),
         label = ifelse(Selected, str_remove(ID, ".jpg"), NA),
         Type = ifelse(Category %in% c("Female", "Male"), "Erotic", "Non-erotic")) |>
  pivot_longer(cols = c("Men_Valence", "Men_Arousal", "Women_Valence", "Women_Arousal")) |>
  separate(name, into=c("Target", "Variable")) |>
  pivot_wider(names_from=Variable, values_from=value) |>
  ggplot(aes(x=Valence, y=Arousal, color=Category)) +
  geom_point(aes(shape=Selected), size=6, alpha=0.8) +
  # ggrepel::geom_label_repel(aes(label = label)) +
  scale_shape_manual(values=c("TRUE"=20, "FALSE"=4)) +
  guides(shape = guide_legend(override.aes = list(color="white"))) +
  facet_grid(~Target) +
  theme_abyss()
```

### Final Selection

-   ID: picture category and number
-   V/H: vertical / horizontal
-   Av/Ap: avoidance - approach dimension
-   M: mean
-   SD: standard deviation
-   JPEG_size80:index of overall complexity
-   LABL: luminance in CIE L*a*b color space
-   LABA: the amount of red color CIE L*a*b color space
-   LABB: the amount of the green color CIE L*a*b color space

```{r}
#| code-fold: true
selection <- df |> 
  filter(ID %in% selected)

write.csv(selection, "stimuli_data.csv", row.names=FALSE)
knitr::kable(selection)
```

```{r}
#| code-fold: true
json <- selection |> 
  select(stimulus=ID, Category, Orientation) |> 
  jsonlite::toJSON()

write(paste("var stimuli_list = ", json), "stimuli_list.js")
```

```{r}
#| code-fold: true
path_ero <- "C:/Dropbox/RECHERCHE/docs/Stimuli/NAPS/NAPS_ERO/"
path_naps <- "C:/Dropbox/RECHERCHE/docs/Stimuli/NAPS/NAPS/"

# Remove all current files
unlink("../stimuli/*")

# Copy each file
for(file in selected) {
  if(str_detect(file, "Male|Female")) {
    file.copy(paste0(path_ero, file), "../stimuli/")
  } else {
    file.copy(paste0(path_naps, file), "../stimuli/")
  }
}
```