### 2023-11-30 experiment/questionnaire_selection/BAIT.qmd
add BAIT 2.0 questionnaire

---
title: "Questionnaire Selection"
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
#| code-fold: true
library(tidyverse)
library(easystats)
library(ggrepel)

df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/main/data/data.csv")
```

This document contains an analysis of the factor structure of the BAIT 1.0 questionnaire (GAAIS + questions about CGI) used in Makowski (Fake Face, 2023).

## Item correlation

```{r}
#| code-fold: true
#| fig.width: 10
#| fig.height: 10

data <- df |> 
  select(starts_with("AI_")) |> 
  rename(
         GAAIS_Neg3_Unethical = AI_2_Unethical,
         GAAIS_Neg10_Dangerous = AI_6_Dangerous,
         GAAIS_Pos7_DailyLife = AI_4_DailyLife,
         GAAIS_Pos12_Exciting = AI_8_Exciting,
         GAAIS_Pos14_Applications = AI_9_Applications,
         # BAIT
         BAIT_1_RealisticImages = AI_1_RealisticImages,
         BAIT_2_VideosReal = AI_3_VideosReal,
         BAIT_3_ImitatingReality = AI_5_ImitatingReality,
         BAIT_4_RealisticVideos = AI_7_RealisticVideos,
         BAIT_5_FaceErrors = AI_10_FaceErrors)

correlation::correlation(data) |> 
  summary() |> 
  plot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) 
```

## GAAIS

- General Attitudes towards Artificial Intelligence Scale (GAAIS; Schepman et al., 2020, 2022)

```{r}
data_gaais <- select(data, starts_with("GAAIS"))

plot(parameters::n_factors(data_gaais))

parameters::factor_analysis(data_gaais, n = 2, rotation = "varimax", sort=TRUE)
```

## BAIT

- Beliefs about Artificial Imaging Technology - BAIT (Makowski, *preprint*)

```{r}
data_bait <- select(data, starts_with("BAIT"))

plot(parameters::n_factors(data_bait))

parameters::factor_analysis(data_bait, n = 2, rotation = "varimax", sort=TRUE)
```