### 2024-06-06 analysis/2_emotions.qmd
@AnafNeves I pushed the 2_emotions script
you can go through it and make sure you can re-run its chunks

---
title: "FictionEro - Affective Impact of Fiction"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(glmmTMB)

options(mc.cores = parallel::detectCores(),
        brms.backend = "cmdstanr",
        width = 300)
```

```{r}
dfsub <- read.csv("../data/data_participants.csv")
df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Sample, Language, Mobile, starts_with("Feedback_"), 
           BAIT_Visual, BAIT_Text, AI_Knowledge, SexualActivity, 
           GAAIS_Negative, GAAIS_Positive, Porn, COPS_Frequency_2,
           -Feedback_Comments),
    by = "Participant"
    ) |> 
  datawizard::rescale(select=c("Valence"), range=c(-1, 1), to=c(0, 1)) |> 
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Relevance =  fct_relevel(Relevance, "Relevant", "Irrelevant", "Non-erotic"),
         Sex =  fct_relevel(Sex, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency_2)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AI-Generated" & Realness < 0.5 ~ "True",
           .default = "False"
         )) |> 
  mutate(across(starts_with("Feedback_"), function(x) {fct_relevel(x, "False", "True")}))

results_table <- function(model, effects="fixed", filter=NULL) {
  if("marginaleffects" %in% class(model)) {
    model |> 
      parameters::parameters() |> 
      as.data.frame() |> 
      select(-Parameter, -SE, -S, z=Statistic, -Predicted) |> 
      insight::format_table() |> 
      parameters::display()
  } else {
    display(parameters::parameters(model, effects=effects, keep=filter))
  }
}
```

## Manipulation Check

This section is about selecting the best model and performing the **manipulation check**: Neutral vs. Emotional

::: {.panel-tabset}

### Arousal

```{r}
df |> 
  mutate(Arousal_Extreme = ifelse(Arousal %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Arousal, fill=Relevance)) +
  geom_histogram(aes(color=Arousal_Extreme), position = "identity", bins=80) +
  facet_grid(Relevance~Sex, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Arousal") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

The *Arousal* data is composed of a mixture of extreme responses (0 and 1) and non-extreme responses between 0 and 1, lending itself to ordered beta model [(Kubinec, 2022)](https://doi.org/10.1017/pan.2022.20), designed explicitly for data with upper and lower bounds, such as survey slider scales. It is a more parsimonious evolution of the zero-and-one inflated beta (ZOIB) regression model (see [this tutorial](https://www.andrewheiss.com/blog/2021/11/08/beta-regression-guide/)) that jointly models the probability of extreme responses, the probability of 1 vs. 0, and the mean and spread of the rest of the responses.

```{r}
#| eval: false

# Get a sense of Random Factors
m1 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1|Participant) + (1|Item), data=df)
m2 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1+Relevance|Participant) + (1|Item), data=df)
m3 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1|Participant) + (1+Sex|Item), data=df)
m4 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1|Participant) + (1+Relevance|Item), data=df)
m5 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1|Participant) + (1+Sex / Relevance|Item), data=df)
m6 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1+Relevance|Participant) + (1+Sex|Item), data=df)
m7 <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (1+Relevance|Participant) + (1+Relevance|Item), data=df)

rbind(
  mutate(performance::icc(m1, by_group=TRUE), Model = "m1"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m1)$R2_conditional, Model = "m1"),
  mutate(performance::icc(m2, by_group=TRUE), Model = "m2"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m2)$R2_conditional, Model = "m2"),
  mutate(performance::icc(m3, by_group=TRUE), Model = "m3"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m3)$R2_conditional, Model = "m3"),
  mutate(performance::icc(m4, by_group=TRUE), Model = "m4"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m4)$R2_conditional, Model = "m4"),
  mutate(performance::icc(m5, by_group=TRUE), Model = "m5"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m5)$R2_conditional, Model = "m5"),
  mutate(performance::icc(m6, by_group=TRUE), Model = "m6"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m6)$R2_conditional, Model = "m6"),
  mutate(performance::icc(m7, by_group=TRUE), Model = "m7"),
  data.frame(Group = "Conditional R2", ICC = performance::r2(m7)$R2_conditional, Model = "m7")
) |>
  ggplot(aes(x = Model, y = ICC, color=Group)) +
  geom_line(aes(group=Group))

test_bf(m6, m1, m2, m3, m4, m5, m7)
compare_performance(m2, m7)

# It seems like the full model, with the effect of relevance on participants and items, is the best.
```

```{r}
m <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance + (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))
```

```{r}
results_table(m)

estimate_relation(m) |> 
  ggplot(aes(x=Relevance, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Relevance), position=position_dodge(width=0.5)) +
  scale_color_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  labs(y="Arousal") +
  facet_wrap(~Sex) +
  theme_abyss()
```

### Enticement

```{r}
df |> 
  mutate(Enticement_Extreme = ifelse(Enticement %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Enticement, fill=Relevance)) +
  geom_histogram(aes(color=Enticement_Extreme), position = "identity", bins=80) +
  facet_grid(Relevance~Sex, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Enticement") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

```{r}
m <- glmmTMB::glmmTMB(Enticement ~ Sex / Relevance + (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
```

```{r}
results_table(m)

estimate_relation(m) |> 
  ggplot(aes(x=Relevance, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Relevance), position=position_dodge(width=0.5)) +
  scale_color_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  labs(y="Enticement") +
  facet_wrap(~Sex) +
  theme_abyss()
```

### Valence

```{r}
df |> 
  mutate(Valence_Extreme = ifelse(Valence %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Valence, fill=Relevance)) +
  geom_histogram(aes(color=Valence_Extreme), position = "identity", bins=80) +
  facet_grid(Relevance~Sex, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Valence") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

```{r}
m <- glmmTMB::glmmTMB(Valence ~ Sex / Relevance + (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
```

```{r}
results_table(m)

estimate_relation(m) |> 
  ggplot(aes(x=Relevance, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Relevance), position=position_dodge(width=0.5)) +
  scale_color_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  labs(y="Enticement") +
  facet_wrap(~Sex) +
  theme_abyss()
```

:::

## Arousal

```{r}
#| fig-height: 12
#| fig-width: 7

plot_descriptive <- function(df, target="Arousal") {
  # By item
  order <- df |>
    mutate(.Item_Order = paste0(Relevance, "_", Sex, "_", Item)) |>
    summarize(mean = mean(.data[[target]], na.rm=TRUE), .by=c("Condition", ".Item_Order")) |> 
    arrange(.Item_Order, desc(Condition)) |> 
    mutate(diff = sign(mean - lag(mean)),
           Relevance = str_extract(.Item_Order, "^[^_]+"),
           .Item_Order = str_extract(.Item_Order, "(?<=_).+")) |>
    filter(Condition == "Photograph") |> 
    arrange(Relevance, diff, mean)

  p1 <- df |>
    mutate(.Item_Order = paste0(Sex, "_", Item)) |>
    group_by(.Item_Order, Condition) |>
    ggdist::mean_qi(.data[[target]], .width=0.95) |>
    mutate(Sex = fct_rev(str_extract(.Item_Order, "^[^_]+"))) |>
    left_join(order[c(".Item_Order", "diff")], by=".Item_Order") |>
    mutate(.Item_Order = fct_relevel(.Item_Order, order$.Item_Order)) |>
    # mutate(.average = mean(.data[[target]]), .by="Item") |>
    # mutate(Item = fct_reorder(Item, .average)) |>
    ggplot(aes(y = .Item_Order, x=.data[[target]])) +
    geom_line(aes(color=as.factor(diff)), linewidth=1) +
    geom_point(aes(color=Condition), size=2) +

    # ggdist::stat_slabinterval(aes(fill=Condition, color=Condition),
    #                           .width=0.95, normalize="xy", slab_alpha=0.5) +
    # ggdist::stat_histinterval(aes(fill=Condition, color=Condition),
    #                           .width=0.95, normalize="xy", slab_alpha=0.5) +
    scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336", "1"="#F44336", "-1"="#2196F3"),
                       breaks = c('AI-Generated', 'Photograph')) +
    # scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
    scale_y_discrete(labels = function(x) str_extract(x, "(?<=_).+")) +
    theme_minimal() +
    theme(panel.grid.minor.x = element_blank(),
          # panel.grid.major.x = element_blank(),
          panel.border = element_blank(),
          axis.line.x = element_blank()) +
    labs(title = paste0(target, " by Item"), color="Condition", y="Items") +
    facet_wrap(~Sex, scales="free_y")

  # By participant
  p2 <- df |>
    summarize(.average = mean(.data[[target]]),
              .by=c("Sex", "Participant", "Condition")) |>
    pivot_wider(names_from = Condition, values_from = .average) |>
    mutate(.diff = Photograph - `AI-Generated`) |>
    mutate(Participant = fct_reorder(Participant, .diff)) |>
    ggplot(aes(x = .diff, y=Participant)) +
    geom_bar(aes(fill=.diff), stat = "identity")  +
    scale_fill_gradient2(guide="none", low="red", high="red", mid="blue") +
    labs(title = paste0(target, " by Participant")) +
    facet_wrap(~Sex, ncol=2, scales="free_y") +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(),
          panel.border = element_blank(),
          axis.line.x = element_blank(),
          axis.text.y = element_blank()) +
    labs(title = paste0(target, " by Participant"), x="Difference (Photograph - AI-Generated)", y="Participants")

  p1 / p2 + plot_layout(heights = c(2, 1))
}

plot_descriptive(df, "Arousal")
```

### Whole Sample

```{r}
m_arousal <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance / Condition + 
                                (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m_arousal)

marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Sex", re.form=NA) |> 
  results_table()
```

### Subsamples

::: {.panel-tabset}

#### Sample

```{r}
m <- glmmTMB::glmmTMB(Arousal ~ Sample / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

#### Mobile

```{r}
m <- glmmTMB::glmmTMB(Arousal ~ Mobile / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

:::

### Moderators

```{r}
test_moderator <- function(df, outcome="Arousal", moderator="BAIT_Visual") {
  f <- paste0(outcome, " ~ Sex / (Condition * ", moderator, ") + (Condition | Participant) + (1|Item)")
  r <- glmmTMB::glmmTMB(as.formula(f),
                      data=filter(df, Relevance == "Relevant"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8)) |>
    parameters::parameters() |> 
    as.data.frame() |> 
    select(-df_error, -Effects, -Group, -Component, -SE)
  r$Moderator <- moderator
  r$Outcome <- outcome
  insight::format_table(r[c(7, 8), ])
}

moderator_results <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Arousal", m))
}

display(moderator_results)
```

```{r}
feedback_results <- data.frame()
for(m in c("Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Arousal", m))
}
display(feedback_results)
```

### Believed

```{r}
m <- glmmTMB::glmmTMB(Arousal ~ Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

moderator_believed <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num", 
           "Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Arousal", m))
}

display(filter(moderator_believed, p < .1))
```

## Enticement

```{r}
#| fig-height: 12
#| fig-width: 7

plot_descriptive(df, "Enticement")
```

### Whole Sample

```{r}
m_enticement <- glmmTMB::glmmTMB(Enticement ~ Sex / Relevance / Condition + 
                                   (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m_enticement)

marginaleffects::avg_slopes(m_enticement, variables="Condition", by="Sex", re.form=NA) |> 
  results_table()
```

### Subsamples

::: {.panel-tabset}

#### Sample

```{r}
m <- glmmTMB::glmmTMB(Enticement ~ Sample / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

#### Mobile

```{r}
m <- glmmTMB::glmmTMB(Enticement ~ Mobile / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

:::

### Moderators

```{r}
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Enticement", m))
}

display(filter(moderator_results, Outcome == "Enticement"))
```

```{r}
for(m in c("Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Enticement", m))
}
display(filter(feedback_results, Outcome == "Enticement"))
```

```{r}
m <- glmmTMB::glmmTMB(Arousal ~ Sex / (Condition * SexualActivity_num) + (Condition | Participant) + (1|Item),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

estimate_relation(m, length=20) |> 
  ggplot(aes(x=SexualActivity_num, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Condition), alpha=0.2) +
  geom_line(aes(color=Condition)) +
  facet_grid(~Sex)
```

### Believed

```{r}
m <- glmmTMB::glmmTMB(Enticement ~ Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

moderator_believed <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num", 
           "Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Enticement", m))
}

display(filter(moderator_believed, p < .1))
```

## Valence

```{r}
#| fig-height: 12
#| fig-width: 7

plot_descriptive(df, "Valence")
```

### Whole Sample

```{r}
m_valence <- glmmTMB::glmmTMB(Valence ~ Sex / Relevance / Condition + 
                                (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m_valence)

marginaleffects::avg_slopes(m_valence, variables="Condition", by="Sex", re.form=NA) |> 
  results_table()
```

### Subsamples

::: {.panel-tabset}

#### Sample

```{r}
m <- glmmTMB::glmmTMB(Valence ~ Sample / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

#### Mobile

```{r}
m <- glmmTMB::glmmTMB(Valence ~ Mobile / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

:::

### Moderators

```{r}
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Valence", m))
}

display(filter(moderator_results, Outcome == "Valence"))
```

```{r}
for(m in c("Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Valence", m))
}
display(filter(feedback_results, Outcome == "Valence"))
```

### Believed

```{r}
m <- glmmTMB::glmmTMB(Valence ~ Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

moderator_believed <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num", 
           "Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Valence", m))
}

display(filter(moderator_believed, p < .1))
```

## Emotionality (Combined Factor)

```{r}
pca <- principal_components(df[c("Valence", "Arousal", "Enticement")], n = 2)
pca
df$Emotionality <- normalize(predict(pca)$PC1)

df |> 
  mutate(Emotionality_Extreme = ifelse(Emotionality %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Emotionality, fill=Relevance)) +
  geom_histogram(aes(color=Emotionality_Extreme), position = "identity", bins=80) +
  facet_grid(Relevance~Sex, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Emotionality") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

```{r}
#| fig-height: 12
#| fig-width: 7

plot_descriptive(df, "Emotionality")
```

### Whole Sample

```{r}
m <- glmmTMB::glmmTMB(Emotionality ~ Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

marginaleffects::avg_slopes(m, variables="Condition", by="Sex", re.form=NA) |> 
  results_table()
```

### Subsamples

::: {.panel-tabset}

#### Sample

```{r}
m <- glmmTMB::glmmTMB(Emotionality ~ Sample / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

#### Mobile

```{r}
m <- glmmTMB::glmmTMB(Emotionality ~ Mobile / Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m, filter="Condition")
```

:::

### Moderators

```{r}
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num")) {
  moderator_results <- rbind(moderator_results,
                             test_moderator(df, "Emotionality", m))
}

display(filter(moderator_results, Outcome == "Emotionality"))
```

```{r}
for(m in c("Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  feedback_results <- rbind(feedback_results,
                            test_moderator(df, "Emotionality", m))
}
display(filter(feedback_results, Outcome == "Emotionality"))
```

### Believed

```{r}
m <- glmmTMB::glmmTMB(Emotionality ~ Sex / Relevance / Condition + (Relevance | Participant) + (1|Item),
                      data=filter(df, ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m)

moderator_believed <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative",
           "Porn", "PornFrequency", "SexualActivity_num", 
           "Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator_believed <- rbind(moderator_believed,
                             test_moderator(filter(df, ConditionBelief == "True"), "Emotionality", m))
}

display(filter(moderator_believed, p < .1))
```

## Figures

```{r}
#| fig-height: 7
#| fig-width: 7

pred1 <- rbind(
  marginaleffects::predictions(m_arousal, by=c("Sex", "Relevance", "Condition"), newdata = "marginalmeans", re.form=NA) |>
  as.data.frame() |> 
  mutate(Outcome="Arousal"),
  marginaleffects::predictions(m_enticement, by=c("Sex", "Relevance", "Condition"), newdata = "marginalmeans", re.form=NA) |>
  as.data.frame() |> 
  mutate(Outcome="Enticement"),
  marginaleffects::predictions(m_valence, by=c("Sex", "Relevance", "Condition"), newdata = "marginalmeans", re.form=NA) |>
  as.data.frame() |> 
  mutate(Outcome="Valence")
)

stars1 <- data.frame(label=c("***", "***", "*", "***"), 
                    Sex=fct_relevel(c("Male", "Male", "Male", "Male"), "Male"), 
                    Relevance=c("Relevant", "Relevant", "Non-erotic", "Relevant"), 
                    Outcome=c("Arousal", "Enticement", "Enticement", "Valence"),
                    y=c(0.75, 0.8, 0.2, 0.8))

p1 <- pred1 |>
  mutate(Sex = fct_relevel(Sex, "Male", "Female")) |> 
  ggplot(aes(x=Relevance, y=estimate)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2), alpha=0.8, linewidth=0.2) +
  geom_pointrange(aes(ymin = conf.low, ymax=conf.high, color=Condition), position = position_dodge(width=0.2), size=0.2) +
  geom_text(data=stars1, aes(y=y, label=label), hjust=0.5, size=3) +
  geom_rect(data=data.frame(Type="All trials", estimate=0.1, Sex=as.factor("Male"), Outcome="Arousal", Relevance="Relevant"), 
            aes(xmin=0.8, xmax=1.2, ymin=0.5, ymax=0.80), color="#8BC34A", alpha=0) +
  geom_rect(data=data.frame(Type="All trials", estimate=0.1, Sex=as.factor("Female"), Outcome="Arousal", Relevance="Relevant"), 
            aes(xmin=0.8, xmax=1.2, ymin=0.2, ymax=0.5), color="#8BC34A", alpha=0) +
  facet_grid(Outcome~Sex, switch="y") +
  scale_y_continuous(limits=c(0, 1), labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1.1, size=9, color="black"),
        legend.text = element_text(size = 10),
        legend.position = ) +
  labs(color="Images presented as:",
       tag = "A")
p1
```

```{r}
m1 <- glmmTMB::glmmTMB(Arousal ~ Sex / Condition + (Condition | Participant) + (1|Item),
                      data=filter(df, Relevance=="Relevant"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m2 <- glmmTMB::glmmTMB(Arousal ~ Sex / Condition + (Condition | Participant) + (1|Item),
                      data=filter(df, Relevance=="Relevant", ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
m3 <- glmmTMB::glmmTMB(Arousal ~ Sex / Condition  + (Condition | Participant) + (1|Item),
                      data=filter(df, Relevance=="Relevant", ConditionBelief == "True", Feedback_AILessArousing=="True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

stars2 <- data.frame(label=c("***", "***", "***", "***", "*"), 
                    Sex=fct_relevel(c("Male", "Male", "Male", "Female", "Female"), 
                                    "Male", "Female"), 
                    Type=c("All trials", "Believed trials", "'AI was less arousing'", "Believed trials", "'AI was less arousing'"), 
                    y=c(0.76, 0.76, 0.76, 0.45, 0.45))

pred2 <- rbind(
  estimate_relation(m1) |> 
    mutate(Type = "All trials"),
  estimate_relation(m2) |> 
    mutate(Type = "Believed trials"),
  estimate_relation(m3) |> 
    mutate(Type = "'AI was less arousing'")
)

p2 <- pred2 |> 
  mutate(Type = fct_relevel(Type, "All trials", "Believed trials", "'AI was less arousing'")) |> 
  ggplot(aes(x=Type, y=Predicted)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2), alpha=0.8, linewidth=0.1) +
  geom_bar(stat="identity", aes(fill=Condition), position = position_dodge(width=0.2), width=0.2, alpha=0.6) +
  geom_pointrange(aes(ymin = CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.2) +
  geom_rect(data=data.frame(Type="All trials", Predicted=0.1, Sex=as.factor("Male")), 
            aes(xmin=0.7, xmax=1.3, ymin=0.55, ymax=0.7), color="#8BC34A", alpha=0) +
  geom_rect(data=data.frame(Type="All trials", Predicted=0.1, Sex=as.factor("Female")), 
            aes(xmin=0.7, xmax=1.3, ymin=0.25, ymax=0.4), color="#8BC34A", alpha=0) +
  geom_text(data=stars2, aes(y=y, label=label), hjust=0.5, size=3) +
  facet_grid(~Sex) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  labs(y="Arousal",
       tag = "B") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=25, hjust=1, vjust=1.1, size=9),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        legend.position = "none")

p1 | p2
```

```{r}
#| fig-width: 10
#| fig-height: 8

m6 <- glm(Feedback_AILessArousing ~ Sex / BAIT_Visual, 
    data=mutate(dfsub, Feedback_AILessArousing = ifelse(Feedback_AILessArousing=="True", 1, 0)), 
    family="binomial")

results_table(m6)

pred3 <- estimate_relation(m6, length=50)

p3 <- pred3 |> 
  mutate(Sex = fct_relevel(Sex, "Male", "Female")) |> 
  ggplot(aes(x=BAIT_Visual, y=Predicted)) +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.3) + 
  geom_line() +
  geom_text(data=data.frame(Sex = as.factor("Male")), aes(x=0.65, y=0.25), label="**") +
  facet_grid(~Sex) +
  scale_y_continuous(labels=scales::percent, expand=c(0, 0)) +
  scale_x_continuous(breaks = c(0.3, 0.5, 0.7, 0.9)) +
  labs(y = "\nProbability of 'AI was less arousing' Feedback\n",
       x = "\nBeliefs about Artificial Information Technology (BAIT)\n'AI algorithms can generate realistic images'",
       tag = "C") +
  theme_minimal() +
  theme(strip.text = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10))

p4 <- patchwork::wrap_plots(p2 / p3 + plot_layout(heights = c(1, 1.6)))
p <- patchwork::wrap_plots(p1 | p4, guides="collect")
p
ggsave("figures/figure1.png", p, width=10, height=8, dpi=300)
```

```{r}
#| eval: false

m4 <- glmmTMB::glmmTMB(Arousal ~ SexualActivity / Sex / Condition + (Condition | Participant) + (1|Item),
                      data=filter(df, Relevance=="Relevant", ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m4)

pred3 <- estimate_relation(m4) 
pred3 |> 
  mutate(xnum = as.numeric(as.factor(SexualActivity))) |>
  ggplot(aes(x=SexualActivity, y=Predicted)) +
  # geom_ribbon(aes(x=xnum, ymin=CI_low, ymax=CI_high, fill=Condition), alpha=0.5) +
  geom_line(aes(group=Condition, color=Condition), position = position_dodge(width=0.2)) +
  geom_pointrange(aes(ymin = CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2)) +
  facet_grid(~Sex)

m5 <- glmmTMB::glmmTMB(Arousal ~ Sex / (Condition * GAAIS_Negative) + (Condition | Participant) + (1|Item),
                      data=filter(df, Relevance=="Relevant", ConditionBelief == "True"), 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
summary(m5)

pred4 <- estimate_relation(m5, length=30) 
pred4 |> 
  ggplot(aes(x=GAAIS_Negative, y=Predicted)) +
  geom_line(aes(group=Condition, color=Condition)) +
  facet_grid(~Sex) 
```

<!-- ## Realness -->

<!-- ### Physical Characteristics -->

<!-- ```{r} -->
<!-- m <- glmmTMB::glmmTMB(Realness ~ Entropy + Complexity + Contrast + Luminance + (1|Participant) + (1|Item), data=df) -->
<!-- results_table(m) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- m <- glmmTMB::glmmTMB(Realness ~ Sex / (Relevance / Condition) + (1|Participant) + (1|Item), data=df, family=beta_family(link = "logit")) -->
<!-- results_table(m) -->
<!-- ``` -->