### 2024-06-06 analysis/3_realness.qmd
part 2/2 of the analysis (@AnafNeves)

---
title: "FictionEro - Determinants of Reality Beliefs"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(glmmTMB)

options(mc.cores = parallel::detectCores(),
        brms.backend = "cmdstanr",
        width = 300)
```

```{r}
dfsub <- read.csv("../data/data_participants.csv")
df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Sample, Language, Mobile, starts_with("Feedback_"), 
           BAIT_Visual, BAIT_Text, AI_Knowledge, SexualActivity, 
           GAAIS_Negative, GAAIS_Positive, Porn, COPS_Frequency_2,
           -Feedback_Comments),
    by = "Participant"
    ) |> 
  datawizard::rescale(select=c("Valence"), range=c(-1, 1), to=c(0, 1)) |> 
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Relevance =  fct_relevel(Relevance, "Relevant", "Irrelevant", "Non-erotic"),
         Sex =  fct_relevel(Sex, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency_2)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AI-Generated" & Realness < 0.5 ~ "True",
           .default = "False"
         ),
         RealnessBelief = ifelse(Realness > 0.5, 1, 0)) |> 
  mutate(across(starts_with("Feedback_"), function(x) {fct_relevel(x, "False", "True")}))

results_table <- function(model, effects="fixed", filter=NULL) {
  if("marginaleffects" %in% class(model)) {
    model |> 
      parameters::parameters() |> 
      as.data.frame() |> 
      select(-Parameter, -SE, -S, z=Statistic, -Predicted) |> 
      insight::format_table() |> 
      parameters::display()
  } else {
    display(parameters::parameters(model, effects=effects, keep=filter))
  }
}
```

### Distribution

```{r}
df |> 
  mutate(Realness_Extreme = ifelse(Realness %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Realness, fill=Relevance)) +
  geom_histogram(aes(color=Realness_Extreme), position = "identity", bins=80) +
  facet_grid(Relevance~Sex, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Relevant"="#E91E63", "Irrelevant"="#FF9800", "Non-erotic"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Realness") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

## Physical Characteristics

```{r}
make_model <- function(f, family=glmmTMB::ordbeta()) {
  m <- glmmTMB::glmmTMB(as.formula(f), 
                        data=df, 
                        family=family,
                        control = glmmTMB::glmmTMBControl(parallel = 8))
  print(results_table(m))
  m
}

m <- make_model("Realness ~ Entropy + (1|Participant) + (1|Item)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Entropy + (1|Participant) + (1|Item)", family="binomial")
m <- make_model("Realness ~ Complexity + (1|Participant) + (1|Item)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Complexity + (1|Participant) + (1|Item)", family="binomial")
m <- make_model("Realness ~ Contrast + (1|Participant) + (1|Item)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Contrast + (1|Participant) + (1|Item)", family="binomial")
m <- make_model("Realness ~ Luminance + (1|Participant) + (1|Item)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Luminance + (1|Participant) + (1|Item)", family="binomial")
```

## Manipulation Check

```{r}
m <- make_model("Realness ~ Sex / (Relevance / Condition) + 
                (Relevance|Participant) + (1|Item)",
                family=glmmTMB::ordbeta())

predbeta <- estimate_relation(m, length=20)

m <- make_model("RealnessBelief ~ Sex / (Relevance / Condition) + 
                (Relevance|Participant) + (1|Item)",
                family="binomial")

predbin <- estimate_relation(m, length=20)

predbeta |> 
  ggplot(aes(x=Relevance, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2)) +
  facet_grid(~Sex)
```

## Moderators

### AI-Attitutes

```{r}
test_moderator <- function(df, moderator="BAIT_Visual") {
  f <- paste0("Realness ~ Sex / (Relevance / Condition * ", moderator, ") + (Relevance | Participant) + (1|Item)")
  r <- glmmTMB::glmmTMB(as.formula(f),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8)) |>
    parameters::parameters() |> 
    as.data.frame() |> 
    select(-df_error, -Effects, -Group, -Component, -SE)
  r$Moderator <- moderator
  insight::format_table(filter(r, str_detect(Parameter, moderator)))
}

moderator <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "AI_Knowledge",
           "GAAIS_Positive", "GAAIS_Negative")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

```{r}
m1 <- glmmTMB::glmmTMB(Realness ~ Sex /
                        (Relevance / Condition *  BAIT_Visual) + 
                        (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m1)

stars1 <- data.frame(Sex = as.factor(c("Male", "Male", "Female", "Female")), 
                    Condition=c("Photograph", "AI-Generated", "Photograph", "AI-Generated"), 
                    label=c("***", "***", "***", "***"),
                    x=c(0.6, 0.6, 0.65, 0.65),
                    y=c(0.8, 0.6, 0.72, 0.55))

pred1 <- estimate_relation(m1)
p1 <- pred1 |> 
  filter(Relevance == "Relevant") |>
  ggplot(aes(x=BAIT_Visual, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Condition), alpha=0.3) +
  geom_line(aes(color=Condition), linewidth=1, key_glyph = draw_key_rect) +
  geom_text(data=stars1, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, size=3) +
  geom_text(data=data.frame(Sex=as.factor("Female"), x=0.65, y=0.638, label="**"), aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  facet_grid(~Sex) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336"), guide="none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\nBeliefs about Artificial Information Technology (BAIT)\n'AI algorithms can generate realistic images'",
       tag = "A")
p1
```

### Porn Usage

```{r}
moderator <- data.frame()
for(m in c("Porn", "PornFrequency", "SexualActivity_num")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

### Feedback

```{r}
moderator <- data.frame()
for(m in c("Feedback_CouldDiscriminate", "Feedback_CouldNotDiscriminate",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

```{r}
m2 <- glmmTMB::glmmTMB(Realness ~ Sex /
                        (Relevance / Feedback_CouldNotDiscriminate / Condition) + 
                        (Relevance | Participant) + (1|Item),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m2)

stars2 <- data.frame(Sex = as.factor(c("Male", "Male", "Female", "Female")), 
                    Condition=c("Photograph", "AI-Generated", "Photograph", "AI-Generated"), 
                    label=c("***", "***", "***", "***"),
                    x=c(1.4, 1.6, 1.4, 1.6),
                    y=c(0.79, 0.72, 0.69, 0.63))
stars3 <- data.frame(Sex = as.factor(c("Male", "Male", "Female")), 
                     label=c("***", "***", "***"),
                     x=c(1.1, 1.8, 1.15),
                     y=c(0.71, 0.80, 0.615))

pred2 <- estimate_relation(m2)

p2 <- pred2 |> 
  filter(Relevance == "Relevant") |>
  ggplot(aes(x=Feedback_CouldNotDiscriminate, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.2), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.8, linewidth=1, key_glyph = "blank") +
  geom_text(data=stars2, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  geom_text(data=stars3, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Sex) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'Could not discriminate AI-generated and real' Feedback",
       tag = "B")
p2
```

## Figures

```{r}
p <- (p1 + theme(legend.position = "none")) / p2  + plot_layout(guides = "collect")
ggsave("figures/figure2.png", p, width=8, height=8, dpi=300)
p
```