### 2024-10-03 study2/experiment/stimuli_selection/selection.qmd
stimuli 2

---
title: "Stimuli Selection "
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
#| code-fold: true

library(tidyverse)
library(easystats)
library(ggrepel)

napsero <- readxl::read_excel("norms_napsero.XLSX", sheet = "Ratings") |>
  full_join(readxl::read_excel("norms_napsero.XLSX", sheet = "Image parameters"), by = join_by("ID", "Category")) |>
  select(-contains("_SD")) |>
  rename(Orientation = `V/H`,
         Hetero_Men_Valence = val_M_HeM,
         Hetero_Women_Valence = val_M_HeF,
         Homo_Men_Valence = val_M_HoM,
         Homo_Women_Valence = val_M_HoF,
         Hetero_Men_Arousal = aro_M_HeM,
         Hetero_Women_Arousal = aro_M_HeF,
         Homo_Men_Arousal = aro_M_HoM,
         Homo_Women_Arousal = aro_M_HoF) |>
  mutate(across(c("Luminance", "Contrast", "JPEG_size80", "LABL", "LABA", "LABB", "Entropy"), as.numeric),
         Description = NA)
```

## Selection

### Select based on norms

In this study, we select only the 25 most arousing items that align with the participants' sexual orientation. For example, homosexual men will be shown the 25 most arousing images of males and male couples. While this approach reduces variability in arousal levels, it is more appropriate for our categorical analysis to focus exclusively on high-arousal stimuli.

Additionally, insights from our previous study suggest that introducing neutral images between arousing ones diminishes participants' ability to maintain arousal, given that arousal tends to build over time. Images that evoke aversion can further disrupt arousal, undermining participants' ability to remain aroused immediately afterward.

```{r}
men <- napsero |>
  filter(Category %in% c("Female", "Opposite-sex Couple")) |>
  group_by(Category) |>
  slice_max(Hetero_Men_Arousal, n=25, with_ties = FALSE) |>
  pull(ID)

homo_men <- napsero |>
  filter(Category %in% c("Male", "Male Couple")) |>
  group_by(Category) |>
  slice_max(Homo_Men_Arousal, n=25, with_ties = FALSE) |>
  pull(ID)

women <- napsero |>
  filter(Category %in% c("Male", "Opposite-sex Couple")) |> 
  group_by(Category) |> 
  slice_max(Hetero_Women_Arousal, n=25, with_ties = FALSE) |>
  pull(ID)

homo_women <- napsero |>
  filter(Category %in% c("Female", "Female Couple")) |>
  group_by(Category) |>
  slice_max(Homo_Women_Arousal, n=25, with_ties = FALSE) |>
  pull(ID)

selected_erotic <- c(men, women, homo_men,homo_women)

cat(
  paste0("N (men) = ", length(men), "\nN (homo_men) = ", length(homo_men),  "\nN (women) = ", length(women), "\nN (homo_women) = ", length(homo_women))
)
```

### Visualization

```{r}
#| fig.width: 10
#| fig.height: 8
#| dpi: 300
#| message: false

dat <- napsero |>
  select(ID, Category, ends_with("_Arousal"), ends_with("_Valence")) |> 
  pivot_longer(cols = c("Hetero_Men_Valence", "Hetero_Men_Arousal", "Hetero_Women_Valence", "Hetero_Women_Arousal", "Homo_Men_Valence", "Homo_Men_Arousal", "Homo_Women_Valence", "Homo_Women_Arousal")) |>
  separate(name, into=c("Participant1", "Participant2", "Variable")) |>
  mutate("Target" = paste0(Participant1, "_", Participant2)) |> 
  select(-Participant1, -Participant2) |> 
  pivot_wider(names_from=Variable, values_from=value) |> 
  mutate(Selected = case_when(
    Target == "Hetero_Men" & ID %in% men ~ TRUE,
    Target == "Homo_Men" & ID %in% homo_men ~ TRUE,
    Target == "Hetero_Women" & ID %in% women ~ TRUE,
    Target == "Homo_Women" & ID %in% homo_women ~ TRUE,
    .default = FALSE
  ))

dat |> 
  ggplot(aes(x = Valence, y = Arousal)) +
  geom_point(data=filter(dat, Selected==FALSE),
             aes(shape = Selected, color = Category), size = 4, alpha = 0.3) +
  geom_point(data=filter(dat, Selected==TRUE),
             aes(shape = Selected, color = Category), size = 4, alpha = 0.8) +
  facet_wrap(~Target) +
  ggside::geom_ysidedensity(data=filter(dat, Selected==TRUE), 
                            aes(color = Category),
                            key_glyph = draw_key_blank, alpha = 0.0) +
  ggside::geom_xsidedensity(data=filter(dat, Selected==TRUE), 
                            aes(color = Category), 
                            key_glyph = draw_key_blank, alpha = 0.0)  +
  scale_shape_manual(values = c("TRUE" = 20, "FALSE" = 4)) +
  scale_color_manual(values = c(
    "Opposite-sex Couple" = "#673AB7", "Male Couple" = "#3F51B5", "Female Couple" = "#9C27B0",
    "Female" = "#E91E63", "Male" = "#2196F3", "Non-erotic"="grey")) +
  guides(shape = guide_legend(override.aes = list(color = "white"))) +
  theme_abyss() +
  theme(ggside.panel.grid.major = element_blank(), 
        ggside.axis.text = element_blank(),
        ggside.axis.line = element_blank())
```

## Final Selection

- ID: picture category and number
- V/H: vertical / horizontal
- Av/Ap: avoidance - approach dimension
- M: mean
- SD: standard deviation
- JPEG_size80: index of overall complexity
- LABL: luminance in CIE L*a*b color space
- LABA: the amount of red color CIE L*a*b color space
- LABB: the amount of the green color CIE L*a*b color space
- Entropy: the amount of complexity in gray scale. Images with high entropy are images that are more "noisy" and have a high degree of contrast.

```{r}
#| code-fold: true

selection <- napsero |> 
  filter(ID %in% selected_erotic)

write.csv(selection, "stimuli_data.csv", row.names=FALSE)
knitr::kable(selection)
```

```{r}
#| code-fold: true
json <- selection |> 
  select(stimulus=ID, Category, Orientation) |> 
  jsonlite::toJSON()

write(paste("var stimuli_list = ", json), "stimuli_list.js")
```

<!-- ```{r} -->
<!-- #| code-fold: true -->
<!-- path_ero <- "C:/Dropbox/RECHERCHE/docs/Stimuli/NAPS/NAPS_ERO/" -->
<!-- path_naps <- "C:/Dropbox/RECHERCHE/docs/Stimuli/NAPS/NAPS/" -->

<!-- # Remove all current files -->
<!-- unlink("../stimuli/*") -->

<!-- # Copy each file -->
<!-- for(file in selected) { -->
<!--   if(str_detect(file, "Male|Female")) { -->
<!--     file.copy(paste0(path_ero, file), "../stimuli/") -->
<!--   } else { -->
<!--     file.copy(paste0(path_naps, file), "../stimuli/") -->
<!--   } -->
<!-- } -->
<!-- ``` -->