### 2025-05-02 study2/analysis/1_cleaning.qmd
updated analysis scripts and data

---
title: "FictionEro-2 - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)

df <- read.csv("../data/rawdata_participants.csv")

dftask <- read.csv("../data/rawdata_task.csv") |>
  full_join(
    df[c("Participant", "Gender", "SexualOrientation")], # add Gneder and sexual orientation to this dataset
    by = join_by(Participant)
    )
```

The initial sample consisted of `r report::report_participants(df, age="Age", sex="Sex", education="Education")`.

## Feedback

::: {.panel-tabset}

### Old vs New feedback

Due to a mistake in the study's implementation, some of the feedback was incorrect, e.g., "Some faces were attractive" and "AI-generated faces were less attractive" rather than "Some images were arousing" and "AI-generated images were less arousing".

```{r}

table(df$version)

```

### Proportion

```{r}
df |> 
  select(starts_with("Feedback"), -Feedback_Text, -Feedback_AllRealConfidence, -Feedback_AllFakeConfidence, -Feedback_DiffSubtle, -Feedback_Enjoyment) |>
  pivot_longer(everything(), names_to = "Question", values_to = "Answer") |>
  group_by(Question, Answer) |> 
  summarise(prop = n()/nrow(df), .groups = 'drop') |> 
  complete(Question, Answer, fill = list(prop = 0)) |> 
  filter(Answer == "True") |> 
  mutate(Question = str_remove(Question, "Feedback_"),
         Question = str_replace(Question, "SomeFacesAttractive", "Some faces were attractive"),
         Question = str_replace(Question, "NoFacesAttractive", "No faces were attractive"),
         Question = str_replace(Question, "AILessAttractive", "AI = Less attractive"),
         Question = str_replace(Question, "AIMoreAttractive", "AI = More attractive"),
         Question = str_replace(Question, "SomeImagesArousing", "Some images were arousing"),
         Question = str_replace(Question, "NoImagesArousing", "No images were arousing"),
         Question = str_replace(Question, "AILessArousing", "AI = Less arousing"),
         Question = str_replace(Question, "AIMoreArousing", "AI = More arousing"),
         # Question = str_replace(Question, "DiffSubtle", "Subtle differences between AI images and photos"),
         Question = str_replace(Question, "DiffNone", "No difference between AI images and photos"),
         Question = str_replace(Question, "LabelsIncorrect", "Labels were incorrect"),
         Question = str_replace(Question, "AllReal", "All images were real"),
         Question = str_replace(Question, "AllFake", "All images were AI-generated"),
         Question = str_replace(Question, "DiffObvious", "Obvious differences between AI images and photos"),
         Question = str_replace(Question, "LabelsReversed", "Labels were reversed")) |>
  mutate(Question = fct_reorder(Question, desc(prop))) |> 
  ggplot(aes(x = Question, y = prop)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks(), labels=scales::percent) +
  labs(x="Feedback", y = "Participants", title = "Feedback") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1), angle = 55, hjust = 1),
    axis.title.x = element_blank()
  )
```

### Confidence

```{r}

df |>
  select(Feedback_AllRealConfidence, Feedback_AllFakeConfidence) |>
  pivot_longer(everything(), names_to = "Authenticity", values_to = "Confidence") |>
  drop_na() |>
  mutate(Authenticity = str_remove(Authenticity, "Feedback_"),
         Authenticity = str_replace(Authenticity, "AllRealConfidence", "All images were real"),
         Authenticity = str_replace(Authenticity, "AllFakeConfidence", "All images were fake")) |>
  ggplot(aes(x = Confidence, fill= Authenticity)) +
  geom_bar() +
  scale_x_continuous(n.breaks = 6) +
  labs(x="Confidence (0-5)", y = "Number of Participants", title = "Confidence on the authenticity of the images") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.title.x = element_blank())
```

### Experiment Enjoyment

```{r}
df |> 
  summarise(n = n(), .by=c("Feedback_Enjoyment")) |>
  filter(!is.na(Feedback_Enjoyment)) |> 
  mutate(n = n / sum(n),
         Feedback_Enjoyment = fct_rev(as.factor(Feedback_Enjoyment))) |>
  ggplot(aes(y = n, x = 1, fill = Feedback_Enjoyment)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values=c("darkgreen", "yellowgreen",  "yellow", "orange", "red")) +
  coord_flip() +
  scale_x_continuous(expand=c(0, 0))  +
  scale_y_continuous(labels = scales::percent)  +
  labs(title="Experiment Enjoyment",
       subtitle="Proportion of participants by condition") + 
  guides(fill = guide_legend(reverse=TRUE)) +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "top",
    legend.title = element_blank()) 
```

### Comments

<!-- eventually create table with selected comments -->

```{r}

data.frame(Comments = df$Feedback_Text) |>
  filter(!tolower(Comments) %in% c(NA, "no", "no comments", "nothing", "none", "", "no.", "no. ", "n/a", "n/a - thanks", "thanks")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE)

```

:::

## Exclusion

```{r}
#| echo: true
#| code-fold: true

outliers <- list()
```

### Mobile

```{r}
mobile <- table(df$Mobile)

# outliers$Mobile <- sum(df$Mobile=="True")
```

There were `r sum(df$Mobile=="True")` participants that used a mobile device.

### Experiment Duration

The experiment's median duration is `r insight::format_value(median(df$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(df$Experiment_Duration, ci=0.5))`).

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 60, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 60, 60, Experiment_Duration),
         Group = ifelse(Participant %in% outliers, "Outlier", "ok")) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_point(aes(color = Group, shape = Category)) +
  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
  geom_vline(xintercept = 15, color = "orange", linetype = "dotted") +
  scale_shape_manual(values = c("extra" = 3, ok = 20)) +
  scale_color_manual(values = c("Outlier" = "red", ok = "black"), guide="none") +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "#4CAF50", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participants"
  )  +
  theme_modern() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

```

### BAIT Questionnaire Duration
#
```{r}
#| fig-width: 15
#| fig-height: 15

df |>
  mutate(Participant = fct_reorder(Participant, Bait_Duration),
         Category = ifelse(Bait_Duration > 5, "extra", "ok"),
         Duration = ifelse(Bait_Duration > 5, 5, Bait_Duration),
         Group = ifelse(Participant %in% outliers, "Outlier", "ok")) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_point(aes(color = Group, shape = Category)) +
  geom_vline(xintercept = median(df$Bait_Duration), color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.5, color = "orange", linetype = "dotted") +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  scale_color_manual(values = c("Outlier" = "red", ok = "black"), guide="none") +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "#9C27B0", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Questionnaire Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  theme_modern() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank(),
        axis.ticks.y = element_blank(),
          axis.text.y = element_blank())

outliers$bait_duration <- filter(df, Bait_Duration < 0.5)$Participant
df[df$Participant %in% outliers$bait_duration, str_detect(names(df), "BAIT_\\d+")] <- NA
```

`r length(outliers$bait_duration)` participants completed the questionnaire in less than 30 seconds.

### Response to Erotic Stimuli

```{r}
# Single arousal response (0)
outliers$novariability <- summarize(dftask, n = length(unique(Arousal)), .by="Participant") |> 
  filter(n == 1) |> 
  select(Participant) |> 
  pull()

df <- filter(df, !Participant %in% outliers$novariability)
dftask <- filter(dftask, !Participant %in% outliers$novariability)
```

`r length(outliers$novariability)` participants showed no variation in their arousal response (did not move the scales).

### Response Coherence

```{r}
dat <- dftask |> 
  filter(!Participant %in% outliers) |> 
  summarize(cor_ArVal = cor(Arousal, Valence),
            cor_ArEnt = cor(Arousal, Enticing),
            .by="Participant")

dat |>
  mutate(Participant = fct_reorder(Participant, cor_ArVal)) |>
  pivot_longer(starts_with("cor_")) |>
  mutate(Group = ifelse(Participant %in% outliers, "Outlier", "ok")) |> 
  mutate(name = fct_relevel(name, "cor_ArVal", "cor_ArEnt"),
         name = fct_recode(name, "Arousal - Valence" = "cor_ArVal", "Arousal - Enticement" = "cor_ArEnt")) |>
  ggplot(aes(y = Participant, x = value)) +
  geom_bar(stat =  "identity") +
  # scale_fill_gradient2(low = "#3F51B5", mid = "#FF9800", high = "#4CAF50", midpoint = 0) +
  # scale_fill_manual(values = c("Outlier" = "red", ok = "black"), guide="none") +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  facet_wrap(~name, ncol=3, scales="free_x")

outliers$emo_ent <- sort(as.character(dat[dat$cor_ArEnt < 0, ]$Participant))
outliers$emo_aro <- sort(as.character(dat[dat$cor_ArVal < 0, ]$Participant))
outliers$arousal <- intersect(outliers$emo_aro, outliers$emo_ent)

```

`r length(outliers$arousal)` participant had a negative correlation between arousal and enticement, AND had a negative correlation between arousal and valance.

Only `r sum(df$Gender=="Other")` participants self-reported as neither female nor male, hence they were excluded from further analysis.

```{r}
df<- filter(df, !Gender=="Other")
df <- filter(df, !Participant %in% outliers$arousal)
dftask <- filter(dftask, !Participant %in% outliers$arousal)
```

## Final Sample

### Summary

The final sample includes `r report::report_participants(df, age="Age", sex="Sex", education="Education")`.

::: {.panel-tabset}

### Age and Gender

```{r}
p_age <- df |>
  ggplot(aes(x = Age, fill = Gender)) + 
  geom_histogram(data=df, aes(x = Age, fill=Gender), binwidth = 2) +
  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5, linetype="dashed") +
  scale_fill_viridis_d() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(10, max(df$Age), by = 10)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

p_age
```

### Sexuality

```{r}
df |>
  ggplot(aes(x = Gender)) +
  geom_bar(aes(fill = SexualOrientation)) +
  scale_y_continuous(expand = c(0, 0), breaks = scales::pretty_breaks()) +
  labs(x = "Biological Sex", y = "Number of Participants", title = "Gender and Sexual Orientation", fill = "Sexual Orientation") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
```

#### Stimuli Choice

```{r}

df |> 
  group_by(Gender) |>
  filter(SexualOrientation == "Bisexual") |>
  ggplot(aes(x=Gender)) +
  geom_bar(aes(fill = StimuliChoice)) +
   labs(x = "Gender", y = "Number of Participants", title = "Bisexuals' Choice of Stimuli", fill = "Stimuli Choice") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

```

### Origin

```{r p_demographics_country}
p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```

### Education

```{r}
p_edu <- df |>
  mutate(Education = fct_relevel(Education, "High school", "Bachelor", "Master", "Doctorate", "Other"))|> 
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1), angle = 45, hjust = 1),
    axis.title.x = element_blank()
  )
p_edu
```

### Ethnicity

```{r}
p_eth <- df |>
  filter(!is.na(Ethnicity)) |> 
  ggplot(aes(x = "", fill = Ethnicity)) +
  geom_bar() +
  coord_polar("y") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )

p_eth

```

### System

```{r p_demographics_settings}
df$Screen_Size <- sqrt(df$Screen_Width * df$Screen_Height)

df |> 
  ggplot(aes(x = Screen_Size, fill=Mobile)) +
  geom_histogram(bins = 50, color="black") +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  labs(x =  expression("Screen Size ("~sqrt(Number~of~Pixels)~")"), y = "Number of Participants", title = "Screen Size") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1))
  )
```

:::

## Save

```{r}
#| code-fold: false

# setdiff(unique(df$Participant), unique(dftask$Participant))
write.csv(df, "../data/data_participants.csv", row.names = FALSE)
write.csv(dftask, "../data/data.csv", row.names = FALSE)
```