### 2025-05-06 study2/analysis/3_realness.qmd
finish* 2_emotions qmd
* missing interpretation of believed, a summary and figures

---
title: "FictionEro2 - Determinants of Reality Beliefs"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---