### 2025-05-07 study2/analysis/3_realness.qmd
updated analysis

---
title: "FictionEro2 - Determinants of Reality Beliefs"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(glmmTMB)

options(mc.cores = parallel::detectCores(),
        brms.backend = "cmdstanr",
        width = 300)
```

```{r}

dfsub <- read.csv("../data/data_participants.csv")

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_", "BAIT_")), COPS_Frequency, SexualActivity,-Feedback_Text),
    by = "Participant"
    ) |> 
  datawizard::rescale(select= c("Arousal", "Enticing", "Valence"), range=c(0, 6), to=c(0,1)) |> 
  datawizard::rescale(select= c("Realness"), range=c(-3,3), to=c(0,1)) |>
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AI-Generated" & Realness < 0.5 ~ "True",
           .default = "False"),
         RealnessBelief = ifelse(Realness > 0.5, 1, 0)) |>
  rename(AllRealConfidence = "Feedback_AllRealConfidence",
         AllFakeConfidence = "Feedback_AllFakeConfidence",
         Enjoyment = "Feedback_Enjoyment"
         )|>
  mutate(across(starts_with("Feedback_"), as.factor))

results_table <- function(model, effects="fixed", filter=NULL) {
  if("marginaleffects" %in% class(model)) {
    model |> 
      parameters::parameters() |> 
      as.data.frame() |> 
      select(-Parameter, -SE, -S, z=Statistic, -Predicted) |> 
      insight::format_table() |> 
      parameters::display()
  } else {
    display(parameters::parameters(model, effects=effects, keep=filter))
  }
}

```

**Note**: [These calculations are based on the validation of the BAIT scale](https://realitybending.github.io/FictionEro/study1/analysis/1_cleaning.html)

```{r}

df$BAIT_Videos <- (df$BAIT_VideosRealistic + (1 - df$BAIT_VideosIssues)) / 2

df$BAIT_Visual = (df$BAIT_ImagesRealistic + (1 - df$BAIT_ImagesIssues) + df$BAIT_ImitatingReality + df$BAIT_EnvironmentReal) / 4

df$BAIT_Text = (df$BAIT_TextRealistic + (1 - df$BAIT_TextIssues)) / 2

df$BAIT_Negative <- (df$BAIT_Worry + df$BAIT_Dangerous)/2

df$BAIT_Positive <- (df$BAIT_Exciting + df$BAIT_Benefit)/2
```

### Distribution

The Realness variable captures a distribution of responses ranging from extreme values (0 and 1) to more ambivalent ratings in between. Given the bounded an ordered beta regression model was employed, as recommended for such data structures [(Kubinec, 2022)](https://doi.org/10.1017/pan.2022.20).

Notably, participants’ confidence in their judgments varied depending on whether they believed the image label (e.g., "AI-generated" or "Photograph"). When the label was believed, participants tended to give more extreme realness ratings, suggesting greater confidence in their evaluations. In contrast, non-belief in the label was associated with a higher frequency of intermediate responses, indicating more uncertainty or ambivalence in perceived realness.

```{r}
#| warning: false

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  mutate(Realness_Extreme = ifelse(Realness %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Realness, fill = Gender)) +
  geom_histogram(aes(color=Realness_Extreme), position = "identity", bins=15) +
  facet_grid(Gender ~ ConditionBelief, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Female"="#FF9800", "Male"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  labs(title = "Realness") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

## Physical Characteristics

## Manipulation Check

This model examines the influence gender and condition on ratings of realness, incorporating random intercepts for participants and items.

<!-- Intercept: Male and Photograph -->

For males the realness ratings were lower for images labelled AI-Generated compared to real photographs. For females they also rated images labelled as AI-Generated as less Real, altough this effect was not significant.

```{r}
#| warning: false

m <- make_model("Realness ~ Gender / Condition + 
                (1|Participant) + (1|Stimulus)",
                family=glmmTMB::ordbeta())

predbeta <- estimate_relation(m, length=20)

m <- make_model("RealnessBelief ~ Gender / Condition + 
                (1|Participant) + (1|Stimulus)",
                family="binomial")

predbin <- estimate_relation(m, length=20)

predbeta |> 
  ggplot(aes(x=Condition, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2)) +
  scale_color_manual(values=c("red", "blue"), guide="none") +
  facet_grid(~Gender)
```

## Moderators

Here we look at whether the findings above are moderated by **AI-Generated, Porn frequency and sexual activity or feedback**

### AI-Attitudes

Attitudes towards AI did not significantly moderate the influence of Gender and Condition on Realness ratings

```{r}
#| warning: false

test_moderator <- function(df, moderator="BAIT_Visual") {
  f <- paste0("Realness ~ Gender / (Condition * ", moderator, ") + (1| Participant) + (1|Stimulus)")
  r <- glmmTMB::glmmTMB(as.formula(f),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8)) |>
    parameters::parameters() |> 
    as.data.frame() |> 
    select(-df_error, -Effects, -Group, -Component, -SE)
  r$Moderator <- moderator
  insight::format_table(filter(r, str_detect(Parameter, moderator)))
}

moderator <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI",
           "BAIT_Positive", "BAIT_Negative")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

<!-- This graph depicts differences in realness ratings based on beliefs about AI's visual abilities for both genders when presented with images labeled as photographs or AI-generated. Notably, images labeled as photographs consistently received higher ratings of realness than those labelled as AI-Generated, even at lower levels of belief in AI's visual capacity.  -->

<!-- As beliefs in AI's ability to generate realistic images increase, both genders show higher realness scores. However, for females with strong beliefs in AI's visual capabilities, realness ratings are higher for images labelled as AI-Generated compared to images labelled as photographs.  -->

<!-- **Notably NONE of these results were significant.** -->

<!-- ```{r} -->
<!-- #| warning: false -->

<!-- m1 <- glmmTMB::glmmTMB(Realness ~ Gender / -->
<!--                         (Condition *  BAIT_Visual) +  -->
<!--                         (1| Participant) + (1|Stimulus), -->
<!--                       data=df,  -->
<!--                       family=glmmTMB::ordbeta(), -->
<!--                       control = glmmTMB::glmmTMBControl(parallel = 8)) -->

<!-- summary(m1) -->

<!-- # stars1 <- data.frame(Gender = as.factor(c("Male", "Male", "Female", "Female")),  -->
<!-- #                     Condition=c("Photograph", "AI-Generated", "Photograph", "AI-Generated"),  -->
<!-- #                     label=c("***", "***", "***", "***"), -->
<!-- #                     x=c(0.6, 0.6, 0.65, 0.65), -->
<!-- #                     y=c(0.8, 0.6, 0.72, 0.55)) -->

<!-- pred1 <- estimate_relation(m1) -->
<!-- p1 <- pred1 |> -->
<!--   ggplot(aes(x=BAIT_Visual, y=Predicted)) + -->
<!--   geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Condition), alpha=0.3) + -->
<!--   geom_line(aes(color=Condition), linewidth=1, key_glyph = draw_key_rect) + -->
<!--   # geom_text(data=stars1, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, size=3) + -->
<!--   # geom_text(data=data.frame(Gender=as.factor("Female"), x=0.65, y=0.638, label="**"), aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) + -->
<!--   facet_grid(~Gender) + -->
<!--   scale_y_continuous(labels=scales::percent) + -->
<!--   scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) + -->
<!--   scale_fill_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336"), guide="none") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.y = element_text(size = 8), -->
<!--         strip.placement = "outside", -->
<!--         strip.background.x = element_rect(fill=c("lightgrey"), color=NA), -->
<!--         strip.text.x = element_text(size = 10), -->
<!--         strip.text.y = element_text(size = 10), -->
<!--         axis.text.x = element_text(size=9, color="black"), -->
<!--         legend.text = element_text(size = 10)) + -->
<!--   labs(y="Realness\n", fill="Images presented as:", color="Images presented as:", -->
<!--        x = "\nBeliefs about Artificial Information Technology (BAIT)\n'AI algorithms can generate realistic images'", -->
<!--        tag = "A") -->
<!-- p1 -->
<!-- ``` -->

### Porn + Sexual Activity

This model evaluates how ratings of realness, predicted by sex, condition, and relevance, are moderated by participants' report of how often they engaged in sexual activity and how often they view pornography.

Looking at porn frequency and sexual activity, the influence of Gender and Condition on Realness scores was not moderated by these variables.

```{r}
#| warning: false

moderator <- data.frame()
for(m in c("PornFrequency", "SexualActivity_num")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

### Feedback

This model evaluates how ratings of realness, predicted by sex, condition, and relevance, are moderated by participants' feedback.

For males, the belief that AI-generated images were less arousing significantly increased realness ratings for images labeled as AI-generated (compared to photos).

For male participants, the belief that all images were real also significantly increased realness ratings for photos.

For female participants, the belief that no faces were attractive was associated with significantly higher realness ratings for photographs.

```{r}
#| warning: false

moderator <- data.frame()
for(m in c("Feedback_DiffObvious", "Feedback_DiffNone",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_NoFacesAttractive", "Feedback_SomeFacesAttractive", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed",
           "Feedback_AllReal",
           "Feedback_AllFake")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

```{r}
#| warning: false

m2 <- glmmTMB::glmmTMB(Realness ~ Gender /
                         (Feedback_AILessArousing / Condition) + 
                        (1 | Participant) + (1|Stimulus),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m2)

stars1 <- data.frame(Gender = as.factor("Male"),
                     Condition = "AI-Generated",
                     label = "***",
                     x = 1.5, 
                     y = 0.55)

pred1 <- estimate_relation(m2)

p1 <- pred1 |> 
  ggplot(aes(x=Feedback_AILessArousing, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.2), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.8, linewidth=1, key_glyph = "blank") +
  # geom_text(data=stars2, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  geom_text(data=stars2, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'AI-Images were less arousing' Feedback",
       tag = "B")

```

```{r}
#| warning: false

m2 <- glmmTMB::glmmTMB(Realness ~ Gender /
                         (Feedback_NoFacesAttractive / Condition) + 
                        (1 | Participant) + (1|Stimulus),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m2)

# stars2 <- data.frame(
#   Gender = as.factor("Male"),
#   Condition = "AI-Generated",
#   label = "***",
#   x = 1.5,
#   y = 0.5)

pred2 <- estimate_relation(m2)

p2 <- pred2 |> 
  ggplot(aes(x=Feedback_NoFacesAttractive, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.2), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2), size=0.8, linewidth=1, key_glyph = "blank") +
  # geom_text(data=stars2, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  # geom_text(data=stars2, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'AI-Faces were less attractive' Feedback",
       tag = "B")

p2
```

## Summary