### 2025-06-17 study2/analysis/3_realness.qmd

---
title: "FictionEro2 - Determinants of Reality Beliefs"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---
## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(glmmTMB)

options(mc.cores = parallel::detectCores(),
        brms.backend = "cmdstanr",
        width = 300)
```

```{r}

dfsub <- read.csv("../data/data_participants.csv")

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_", "BAIT_")), COPS_Frequency, SexualActivity,-Feedback_Text),
    by = "Participant"
    ) |> 
  datawizard::rescale(select= c("Arousal", "Enticing", "Valence"), range=c(0, 6), to=c(0,1)) |> 
  datawizard::rescale(select= c("Realness"), range=c(-3,3), to=c(0,1)) |>
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         SexualOrientation = fct_relevel(SexualOrientation,"Heterosexual", "Homosexual", "Bisexual"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AI-Generated" & Realness < 0.5 ~ "True",
           .default = "False"),
         RealnessBelief = ifelse(Realness > 0.5, 1, 0)) |>
  rename(AllRealConfidence = "Feedback_AllRealConfidence",
         AllFakeConfidence = "Feedback_AllFakeConfidence",
         Enjoyment = "Feedback_Enjoyment"
         )|>
  mutate(across(starts_with("Feedback_"), as.factor)) |>
  mutate(StimuliType = case_when(
    grepl("couple", Stimulus, ignore.case = TRUE) ~ "Couple",
    TRUE ~ "Individual")) |>
  mutate(StimuliType = fct_relevel(StimuliType, "Individual", "Couple"))

results_table <- function(model, effects="fixed", filter=NULL) {
  if("marginaleffects" %in% class(model)) {
    model |> 
      parameters::parameters() |> 
      as.data.frame() |> 
      select(-Parameter, -SE, -S, z=Statistic, -Predicted) |> 
      insight::format_table() |> 
      parameters::display()
  } else {
    display(parameters::parameters(model, effects=effects, keep=filter))
  }
}

```

**Note**: [These calculations are based on the validation of the BAIT scale](https://realitybending.github.io/FictionEro/study1/analysis/1_cleaning.html)

```{r}

df$BAIT_Videos <- (df$BAIT_VideosRealistic + (1 - df$BAIT_VideosIssues)) / 2

df$BAIT_Visual = (df$BAIT_ImagesRealistic + (1 - df$BAIT_ImagesIssues) + df$BAIT_ImitatingReality + df$BAIT_EnvironmentReal) / 4

df$BAIT_Text = (df$BAIT_TextRealistic + (1 - df$BAIT_TextIssues)) / 2

df$BAIT_Negative <- (df$BAIT_Worry + df$BAIT_Dangerous)/2

df$BAIT_Positive <- (df$BAIT_Exciting + df$BAIT_Benefit)/2
```

### Distribution

The Realness variable captures a distribution of responses ranging from extreme values (0 and 1) to more ambivalent ratings in between. Given the bounded an ordered beta regression model was employed, as recommended for such data structures [(Kubinec, 2022)](https://doi.org/10.1017/pan.2022.20).

Notably, participants’ confidence in their judgments varied depending on whether they believed the image label (e.g., "AI-generated" or "Photograph"). When the label was believed, participants tended to give more extreme realness ratings, suggesting greater confidence in their evaluations. In contrast, non-belief in the label was associated with a higher frequency of intermediate responses, indicating more uncertainty or ambivalence in perceived realness.

```{r}
#| warning: false

df |> 
  mutate(ConditionBelief = as.factor(ConditionBelief)) |> 
  mutate(ConditionBelief = case_when(
    ConditionBelief == "False" ~ "Not Believed",
    ConditionBelief == "True" ~ "Believed",)) |>
  mutate(Realness_Extreme = ifelse(Realness %in% c(0, 1), "Extreme", "Not extreme")) |>
  ggplot(aes(x=Realness, fill = Gender)) +
  geom_histogram(aes(color=Realness_Extreme), position = "identity", bins=15) +
  scale_x_continuous(
    breaks = c(0, 1),
    labels = c("AI","Real")) +
  facet_grid(Gender ~ ConditionBelief, scales="free_y") +
  scale_color_manual(values=c("red", "white"), guide="none") +
  scale_fill_manual(values=c("Female"="#FF9800", "Male"="#03A9F4"), guide="none") +
  scale_y_sqrt(expand = c(0, 0)) +
  facet_wrap(Gender~SexualOrientation) +
  labs(title = "Realness") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
```

```{r}

df <- df|>
  filter(SexualOrientation == "Heterosexual")
```

## Physical Characteristics

Here, we are examining the effect of various **physical characteristics** of the images to evaluate whether they influence the perceived realness of the images. Neither Entropy, Complexity, Contrast, nor Luminance significantly influence the perceived realness of the images, nor the likelihood of realness being above 0.5 ("RealnessBelief"), as indicated by the lack of significant effects.

```{r}
#| warning: false

make_model <- function(f, family=glmmTMB::ordbeta()) {
  m <- glmmTMB::glmmTMB(as.formula(f), 
                        data=df, 
                        family=family,
                        control = glmmTMB::glmmTMBControl(parallel = 8))
  print(results_table(m))
  m
}

m <- make_model("Realness ~ Entropy + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Entropy + (1|Participant) + (1|Stimulus)", family="binomial")
m <- make_model("Realness ~ Complexity + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Complexity + (1|Participant) + (1|Stimulus)", family="binomial")
m <- make_model("Realness ~ Contrast + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Contrast + (1|Participant) + (1|Stimulus)", family="binomial")
m <- make_model("Realness ~ Luminance + (1|Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m <- make_model("RealnessBelief ~ Luminance + (1|Participant) + (1|Stimulus)", family="binomial")
```

## Manipulation Check

This model examines the influence **Gender and Condition** on ratings of realness, incorporating random intercepts for participants and items.

For males the realness ratings were lower for images labelled AI-Generated compared to real photographs. For females they also rated images labelled as AI-Generated as less Real, altough this effect was not significant.

**Note**. Adding StimuliType as a predictor did not change the pattern - Condition was still only a significant predictor for men.

```{r}
#| warning: false

m <- make_model("Realness ~ Gender /Condition + 
                (1|Participant) + (1|Stimulus)",
                family=glmmTMB::ordbeta())

predbeta <- estimate_relation(m, length=20)

m <- make_model("RealnessBelief ~ Gender /Condition + 
                (1|Participant) + (1|Stimulus)",
                family="binomial")

predbin <- estimate_relation(m, length=20)

predbeta |> 
  ggplot(aes(x=Condition, y=Predicted)) +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.2)) +
  scale_color_manual(values=c("red", "blue"), guide="none") +
  facet_grid(StimuliType~Gender)
```

## Moderators

Here we look at whether the findings above are moderated by **AI-Generated, Porn frequency and sexual activity or feedback**. Note that the focus solely on heterosexaul individuals (N = `r sum(dfsub$SexualOrientation == "Heterosexual")`)

### AI-Attitudes

This model looks at whether realness ratings are moderated by attitudes towards AI.  Looking at the results, the influence of Gender and Condition on Realness scores was not moderated by attidues towards AI.

```{r}
#| warning: false

test_moderator <- function(df, moderator="BAIT_Visual") {
  f <- paste0("Realness ~ Gender / (Condition * ", moderator, ") + (1| Participant) + (1|Stimulus)")
  r <- glmmTMB::glmmTMB(as.formula(f),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8)) |>
    parameters::parameters() |> 
    as.data.frame() |> 
    select(-df_error, -Effects, -Group, -Component, -SE)
  r$Moderator <- moderator
  insight::format_table(filter(r, str_detect(Parameter, moderator)))
}

moderator <- data.frame()
for(m in c("BAIT_Visual", "BAIT_Text", "BAIT_ExpertAI",
           "BAIT_Positive", "BAIT_Negative")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

### Porn + Sexual Activity

This model evaluates how ratings of realness, predicted by Gender and Condition, are moderated by participants' report of how often they engaged in sexual activity and how often they view pornography. Looking at porn frequency and sexual activity, the influence of Gender and Condition on Realness scores was not moderated by these variables.

```{r}
#| warning: false

moderator <- data.frame()
for(m in c("PornFrequency", "SexualActivity_num")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

### Feedback

This model evaluates how ratings of realness, predicted by sex, condition, and relevance, are moderated by participants' feedback.

For males, the feedback that AI-generated images were less arousing significantly moderated realness ratings for images labeled as AI-generated, with higher realness ratings for those images when participants thought they were less arousing.

For males, the belief that all images were real significantly increased realness ratings for all images, regardless of their actual content.

For females, the belief that no faces were attractive significantly increased realness ratings for the images, suggesting that lower attractiveness perceptions made the images seem more real.

```{r}
#| warning: false

moderator <- data.frame()
for(m in c("Feedback_DiffObvious", "Feedback_DiffNone",
           "Feedback_AIMoreArousing", "Feedback_AILessArousing", 
           "Feedback_NoFacesAttractive", "Feedback_SomeFacesAttractive", 
           "Feedback_LabelsIncorrect", "Feedback_LabelsReversed",
           "Feedback_AllReal", "Feedback_AllFake",
           "AllRealConfidence","AllFakeConfidence" ,
           "Enjoyment")) {
  moderator <- rbind(moderator, test_moderator(df, m))
}

display(filter(moderator, p < .05))
```

```{r}
#| warning: false

m2 <- glmmTMB::glmmTMB(Realness ~ Gender /
                         (Feedback_AILessArousing / Condition) + 
                        (1|Participant) + (1|Stimulus),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m2)

stars2 <- data.frame(Gender = as.factor(c("Male")),
                     feedback = as.factor("Feedback_AILessArousing"),
                    label=c("***"),
                    x=c(1.0),
                    y=c(0.545))
# stars3 <- data.frame(Sex = as.factor(c("Male", "Male", "Female")), 
#                      label=c("***", "***", "***"),
#                      x=c(1.1, 1.8, 1.15),
#                      y=c(0.71, 0.80, 0.615))

pred1 <- estimate_relation(m2)

p1 <- pred1 |> 
  ggplot(aes(x=Feedback_AILessArousing, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.3), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.3), size=0.8, linewidth=1, key_glyph = "blank") +
  # geom_text(data=stars2, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  geom_text(data=stars2, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'AI-Images were less arousing' Feedback",
       tag = "A")

p1

```

```{r}
#| warning: false

m2 <- glmmTMB::glmmTMB(Realness ~ Gender /
                         (Feedback_NoFacesAttractive / Condition) + 
                        (1 | Participant) + (1|Stimulus),
                      data=df, 
                      family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))

summary(m2)

stars3 <- data.frame(Gender = as.factor(c("Female", "Female")), 
                    Condition=c("Photograph", "AI-Generated"), 
                    label=c("**", "**"),
                    x=c(1.3, 1.7),
                    y=c(0.6, 0.6))

stars4 <- data.frame(Gender = as.factor("Male"), 
                    Condition="AI-Generated", 
                    label="***",
                    x=c(1),
                    y=c(0.545))

pred2 <- estimate_relation(m2)

p2 <- pred2 |> 
  ggplot(aes(x=Feedback_NoFacesAttractive, y=Predicted)) +
  geom_line(aes(color=Condition, group=Condition), linewidth=.3, position = position_dodge(width=0.4), key_glyph = "path") +
  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Condition), position = position_dodge(width=0.4), size=0.8, linewidth=1, key_glyph = "blank") +
  geom_text(data=stars3, aes(x=x, y=y, label=label, color=Condition), hjust=0.5, key_glyph = "blank", size=3) +
  geom_text(data=stars4, aes(x=x, y=y, label=label), hjust=0.5, color="#424242", size=3) +
  guides(color=guide_legend(override.aes = list(linewidth = 1.5))) +
  facet_grid(~Gender) +
  scale_y_continuous(labels=scales::percent) +
  scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        strip.placement = "outside",
        strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        axis.text.x = element_text(size=9, color="black"),
        legend.text = element_text(size = 10)) +
  labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "\n'No faces attractive' Feedback",
       tag = "B")

p2
```

### Association betwen Emotional responses and Realness Ratings

**Realness** ratings were positively correlated with **Arousal**, **Enticement**, and **Valence.**

To examine the unique role of **Arousal** in predicting **perceived realness**, a model was tested including both linear and quadratic effects of Arousal, along with its interactions with Gender and Condition (AI-generated vs. real images), modeling participants and images as random intercepts. 
To assess whether a non-linear (quadratic) component improved the model, a comparison was conducted between a model including a quadratic term for Arousal and one without. The model excluding the quadratic term was decisively favored, with Bayes Factors indicating that the data were over 1000 times more likely under the simpler linear model. This strongly supports the adequacy of a linear relationship between Arousal and Realness in this context. The following findings were seen in the simpler model:

- **Arousal** had a strong positive linear effect on Realness, indicating that higher arousal was **associated** with **increased perceptions of realness**.

- For **female** participants, the positive **effect of Arousal on Realness** is significantly **weaker** than for **males**. This suggests that women are less influenced by Arousal when making realness judgments.

- Among **males**, the effect of **Arousal, on realness ratings**, is significantly **weaker** when judging images previous labelled as **AI-generated** compared to real ones.

These results suggest a gender-specific modulation in how emotional arousal shapes beliefs about what is real, particularly when distinguishing between real and synthetic faces.

```{r}
df |>
  select(Gender, Arousal, Enticing, Valence, RealnessBelief, Realness) |>
  group_by(Gender) |>
  correlation::correlation()

p_a <- ggplot(df, aes(x = Arousal, y = Realness, color = Condition, fill = Condition)) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~Gender) +
  scale_color_manual(values = c("AI-Generated" = "#2196F3", "Photograph" = "#F44336")) +
  scale_fill_manual(values = c("AI-Generated" = "#2196F3", "Photograph" = "#F44336")) +
  smplot2::sm_statCorr() +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    strip.placement = "outside",
    strip.background.x = element_rect(fill = "lightgrey", color = NA),
    strip.text.x = element_text(size = 10),
    strip.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 9, color = "black"),
    legend.text = element_text(size = 10)) +
   labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "Arousal")

p_e <- ggplot(df, aes(x = Enticing, y = Realness, color = Condition, fill = Condition)) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~Gender) +
  scale_color_manual(values = c("AI-Generated" = "#2196F3", "Photograph" = "#F44336")) +
  scale_fill_manual(values = c("AI-Generated" = "#2196F3", "Photograph" = "#F44336")) +
  smplot2::sm_statCorr() +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    strip.placement = "outside",
    axis.title.y = element_blank(),
    strip.background.x = element_rect(fill = "lightgrey", color = NA),
    strip.text.x = element_text(size = 10),
    axis.text.x = element_text(size = 9, color = "black"),
    legend.text = element_text(size = 10)) +
   labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "Enticement")

p_v <- ggplot(df, aes(x = Valence, y = Realness, color = Condition, fill = Condition)) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~Gender) +
  scale_color_manual(values = c("AI-Generated" = "#2196F3", "Photograph" = "#F44336")) +
  scale_fill_manual(values = c("AI-Generated" = "#2196F3", "Photograph" = "#F44336")) +
  smplot2::sm_statCorr() +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    strip.placement = "outside",
    axis.title.y = element_blank(),
    strip.background.x = element_rect(fill = "lightgrey", color = NA),
    strip.text.x = element_text(size = 10),
    axis.text.x = element_text(size = 9, color = "black"),
    legend.text = element_text(size = 10)) +
   labs(y="Realness\n", fill="Images presented as:", color="Images presented as:",
       x = "Valence")

p_sub <- (p_a + theme(legend.position = "none"))| (p_e + theme(legend.position = "none"))| p_v + plot_layout(guides = "collect")

m1_dir <- make_model("Realness ~ Arousal / Gender/ Condition + (1 | Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m2_dir <- make_model("Realness ~ poly(Arousal, 2) / Gender/ Condition + (1 | Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())

compare_performance(m1_dir, m2_dir)
test_performance(m1_dir, m2_dir)

```

## Summary

- For **males** the **realness ratings were lower for images labelled AI-Generated** compared to real photographs. For females they also rated images labelled as AI-Generated as less Real, altough this effect was not significant. 
- Looking at moderators, **attitudes towards AI**, participants reports of **how often they engaged in sexual activity** and **how often they view pornography**, did **not moderate realness ratings**
- Looking at the **feedback as a moderator** for realness ratings, for **males**, feedback that **AI-generated images were less arousing increased realness ratings for those images**, and believing all images were real raised realness ratings across all images. For females, the **reporting that no faces were attractive increased realness ratings** for the images.

## Figures

```{r}

p <- (p1 + theme(legend.position = "none")) / p2  + plot_layout(guides = "collect")
p
```

## Notes/ Questions

1. 'Other' was removed due to the very small percentage of individuals identifying as such.
2. 'Porn' factor was not computed, instead sexual activity and frequency of watching porn were treated seperate. 
3. Missing significance starts on figures.