### 2025-12-09 analysis/analysis.qmd
Update analysis.qmd

---
title: "FictionEro - Affective Impact of Fiction"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(glmmTMB)
library(gt)
library(ggdist)

options(mc.cores = parallel::detectCores(),
        brms.backend = "cmdstanr",
        width = 300)

results_table <- function(model, effects="fixed", filter=NULL) {
  if("marginaleffects" %in% class(model)) {
    model |> 
      parameters::parameters() |> 
      as.data.frame() |> 
      select(-Parameter, -SE, -S, z=Statistic) |> 
      insight::format_table() |> 
      parameters::display()
  } else {
    display(parameters::parameters(model, effects=effects, keep=filter))
  }
}

params_table <- function(m) {
  d <- as.data.frame(m, variable = insight::find_parameters(m, effects = "fixed", flatten  = TRUE))
  t <- bayestestR::describe_posterior(d, effects = "fixed", component = "all", ci_method = "hdi", test = "p_direction")
  out <- insight::format_table(t)
  out |> 
    mutate(col = t$pd) |> 
    gt::gt() |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.95, palette = "yellow") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.97, palette = "green") |> 
    gt::cols_hide("col")
}
```

```{r data study1}
#| warning: false

dfsub1 <- dfsub <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionEro/refs/heads/main/study1/data/data_participants.csv")
df1 <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionEro/refs/heads/main/study1/data/data.csv") |> 
  right_join(
    select(dfsub1, Participant, Sample, Language, Mobile, starts_with("Feedback_"), 
           BAIT_Visual, BAIT_Text, AI_Knowledge, SexualActivity, 
           GAAIS_Negative, GAAIS_Positive, Porn, COPS_Frequency_2,
           Screen_Size,
           -Feedback_Comments),
    by = "Participant") |> 
  rename(Gender = Sex) |>
  datawizard::rescale(select=c("Valence"), range=c(-1, 1), to=c(0, 1)) |> 
  mutate(Condition = fct_recode(Condition, "Photograph" = "Photograph", "AIGenerated" = "AI-Generated"),
         Condition = fct_relevel(Condition, "Photograph", "AIGenerated"),
         Relevance = fct_recode(Relevance,  "Relevant" = "Relevant",  "Irrelevant"= "Irrelevant", "NonErotic" = "Non-erotic"),
         Relevance =  fct_relevel(Relevance, "Relevant", "Irrelevant", "NonErotic"),
         Gender =  fct_relevel(Gender, "Male", "Female"),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AIGenerated" & Realness < 0.5 ~ "True",
           .default = "False"),
         ConditionBelief = as.factor(ConditionBelief),
         ConditionBelief = fct_relevel(ConditionBelief, "True", "False"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency_2)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         across(starts_with("Feedback_"), function(x) {fct_relevel(x, "False", "True")}))
```

```{r data study2}
#| warning: false

dfsub2 <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionEro/refs/heads/main/study2/data/data_participants.csv")
df2 <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionEro/refs/heads/main/study2/data/data.csv") |> 
  right_join(
    select(dfsub2, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency, SexualActivity,- Feedback_Text),
    by = "Participant"
  ) |> 
  datawizard::rescale(select= c("Arousal", "Enticing", "Valence"), range=c(0, 6), to=c(0,1)) |>
  datawizard::rescale(select= c("Realness"), range=c(-3,3), to=c(0,1)) |>
  filter(SexualOrientation == "Heterosexual") |> # Keep only heterosexuals
  mutate(Gender =  fct_relevel(Gender, "Male", "Female"),
         Condition = case_when(
           Condition == "Fiction" ~ "AIGenerated",
           Condition == "Reality" ~ "Photograph"), 
         Condition = fct_relevel(Condition, "Photograph", "AIGenerated"),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0.5 ~ "True",
           Condition == "AIGenerated" & Realness < 0.5 ~ "True",
           .default = "False"),
         ConditionBelief = as.factor(ConditionBelief),
         ConditionBelief = fct_relevel(ConditionBelief, "True", "False"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency)),
         SexualActivity_num = as.numeric(as.factor(SexualActivity)),
         across(starts_with("Feedback_"), as.factor), 
         across(starts_with("Feedback_"), function(x) {fct_relevel(x, "False", "True")})) |>
    rename(Item = Stimulus,
           Enticement = Enticing,
           AllRealConfidence = "Feedback_AllRealConfidence",
           AllFakeConfidence = "Feedback_AllFakeConfidence",
           Enjoyment = "Feedback_Enjoyment") 
```

## Manipulation Checks

::: {.panel-tabset}

### Rate Belief

```{r ratebelief}

rate1 <- table(df1$Feedback_LabelsIncorrect)["True"] / sum(table(df1$Feedback_LabelsIncorrect))
rate2 <- table(df2$Feedback_AllReal)["True"] / sum(table(df2$Feedback_AllReal))

```

The rate of people that did not believe the instructions in study 1 (i.e., gave feedback that the labels were incorrect) was `r round(mean(rate1) * 100, 1)`% while in study 2 (i.e., gave feedback that all images were real images) was `r round(mean(rate2) * 100, 1)`%.

### Arousal Distributions

```{r Arousal1}
#| fig-width: 12
arousal_dist1 <- df1 |>
  ggplot(aes(x=Arousal, fill = Condition)) +
  geom_histogram(position = "identity", alpha = 0.5, colour = "black", bins = 30) +
  scale_x_continuous(breaks = c(0, 1)) +
  labs(title = "Arousal Study 1", ) +
  facet_wrap(~Gender) +
  scale_fill_manual(values=c("AIGenerated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

arousal_dist1
```

```{r Arousal2}
#| fig-width: 12
arousal_dist2 <- df2 |>
  ggplot(aes(x=Arousal, fill = Condition)) +
  geom_bar(position = "identity", alpha = 0.5, colour = "black") +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(breaks = c(0, 1)) +
  labs(title = "Arousal Study 2", ) +
  facet_wrap(~Gender) +
  scale_fill_manual(values=c("AIGenerated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

arousal_dist2
```

### Enticment Distributions

```{r Arousal1}
#| fig-width: 12
enticement_dist1 <- df1 |>
  ggplot(aes(x=Enticement, fill = Condition)) +
  geom_histogram(position = "identity", alpha = 0.5, colour = "black", bins = 30) +
  scale_x_continuous(breaks = c(0, 1)) +
  labs(title = "Enticement Study 1", ) +
  facet_wrap(~Gender) +
  scale_fill_manual(values=c("AIGenerated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

enticement_dist1
```

```{r Arousal2}
#| fig-width: 12
enticement_dist2 <- df2 |>
  ggplot(aes(x=Enticement, fill = Condition)) +
  geom_bar(position = "identity", alpha = 0.5, colour = "black") +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(breaks = c(0, 1)) +
  labs(title = "Enticment Study 2", ) +
  facet_wrap(~Gender) +
  scale_fill_manual(values=c("Photograph"="#F44336", "AIGenerated"="#2196F3")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

enticement_dist2
```

### Valence Distributions

```{r Arousal1}
#| fig-width: 12
valence_dist1 <- df1 |>
  ggplot(aes(x=Valence, fill = Condition)) +
  geom_histogram(position = "identity", alpha = 0.5, colour = "black", bins = 30) +
  scale_x_continuous(breaks = c(0, 1)) +
  labs(title = "Valence Study 1", ) +
  facet_wrap(~Gender) +
  scale_fill_manual(values=c("AIGenerated"="#2196F3", "Photograph"="#F44336")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

valence_dist1
```

```{r Arousal2}
#| fig-width: 12
valence_dist2 <- df2 |>
  ggplot(aes(x=Valence, fill = Condition)) +
  geom_bar(position = "identity", alpha = 0.5, colour = "black") +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(breaks = c(0, 1)) +
  labs(title = "Valence Study 2", ) +
  facet_wrap(~Gender) +
  scale_fill_manual(values=c("Photograph"="#F44336", "AIGenerated"="#2196F3")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))

valence_dist2
```
:::

## Average Arousal

::: {.panel-tabset}

### Study 1

```{r meanarousal 1}
#| message: false

dfsub1 <- dfsub1 |>
  right_join(
    df1 |>
      group_by(Participant, Condition) |>
      summarise(mean_arousal = mean(Arousal, na.rm = TRUE), .groups = "drop") |>
      pivot_wider(
        names_from = Condition,
        names_prefix = "mean_Arousal_",
        values_from = mean_arousal
      ) |>
      mutate(diff_arousal = `mean_Arousal_AIGenerated` - mean_Arousal_Photograph)
  )

ttest1 <- stats::t.test(diff_arousal ~ Feedback_AILessArousing, data = dfsub1)
```

```{r}
diff_plot1 <- dfsub1 |> 
  ggplot(aes(x = Feedback_AILessArousing, y = diff_arousal, fill = Feedback_AILessArousing)) +
  geom_boxplot(width = 0.4,alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.1, size = 1) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 2, fill = "black") +
  labs(
    title = "Study 1",
    x = "Participant said AI is less arousing",
    y = "Arousal Difference (AI - Photo)"
  ) +
  scale_fill_manual(values = c("True" = "#E66100", "False" = "#5D3A9B")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "none"
  )

diff_plot1
```

### Study 2

```{r meanarousal 2}
#| message: false

dfsub2 <- dfsub2 |>
  right_join(
    df2|>
      group_by(Participant, Condition)|>
      summarise(mean_arousal = mean(Arousal, na.rm = TRUE), .groups = "drop") |>
      pivot_wider(names_from = Condition, 
                  names_prefix = "mean_Arousal_", 
                  values_from = mean_arousal) |>
      mutate(diff_arousal = `mean_Arousal_AIGenerated` - mean_Arousal_Photograph)
  )

ttest2 <- stats::t.test(diff_arousal ~ Feedback_AILessArousing, data = dfsub2)
```

```{r}
diff_plot2 <- dfsub2 |> 
  ggplot(aes(x = Feedback_AILessArousing, y = diff_arousal, fill = Feedback_AILessArousing)) +
  geom_boxplot(width = 0.4, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.1, size = 1) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 2, fill = "black") +
  labs(
    title = "Study 2",
    x = "Participant said AI is less arousing",
    y = "Arousal Difference (AI - Photo)"
  ) +
  scale_fill_manual(values = c("True" = "#E66100", "False" = "#5D3A9B")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "none"
  )

diff_plot2
```

**Study 1**: `r report::report(ttest1, data = dfsub1)`

**Study 2**: `r report::report(ttest2, data = dfsub2)`

:::

## Effect of Manipulation

::: {.panel-tabset}

### Arousal

#### Distributions

```{r}
df1 |> 
  ggplot(aes(x = Gender, y = Arousal, fill = ConditionBelief)) +
  geom_violin(position = position_dodge(width = 1), alpha = 0.7, trim = FALSE) +
  stat_pointinterval(position = position_dodge(width = 1),point_size = 1.5) +
  facet_grid(Relevance ~ Condition) +
  scale_fill_manual(values = c("True" = "#E66100", "False" = "#5D3A9B")) +
  theme_minimal(base_size = 14) +
  labs(title = "Arousal by Gender × Belief × Relevance × Condition", subtitle ="Study 1", x = "Gender",y = "Arousal",fill = " Condition Belief")

df2 |> 
  ggplot(aes(x = Gender, y = Arousal, fill = ConditionBelief)) +
  geom_violin(position = position_dodge(width = 1), alpha = 0.7, trim = FALSE) +
  stat_pointinterval(position = position_dodge(width = 1),point_size = 1.5) +
  facet_grid( ~ Condition) +
  scale_fill_manual(values = c("True" = "#E66100", "False" = "#5D3A9B")) +
  theme_minimal(base_size = 14) +
  labs(title = "Arousal by Gender × Belief × Relevance × Condition", subtitle ="Study 2", x = "Gender", y = "Arousal",fill = " Condition Belief")

```

```{r}
arousal1 <- readRDS("models/ModelArousal_1.rds")
arousal2 <- readRDS("models/ModelArousal_2.rds")
```

#### Posterior Description

<!-- Sampling Diagnostics -->

<!-- - ESS should be at least 1000 (the effective sample size - ESS - is an estimation of the number of independent samples) -->

<!-- - R^(“R-hat”) Should be lower than 1.01 (R-hat is a measure of convergence) -->

<!-- ```{r} -->
<!-- results_table(m_arousal1) -->
<!-- results_table(m_arousal2) -->
<!-- ``` -->

```{r}
params_table(arousal1)
summary(arousal1, robust = TRUE)
```

```{r}
params_table(arousal2) 
summary(arousal2, robust = TRUE)
```

#### Model Performance

```{r}
performance::performance(arousal1)
```

```{r}
performance::performance(arousal2)
```

:::

## Interindividual Variability and Moderators

1. extract random effects
2. make plot of participant effects
3. effect of condition believed + relevant 
4. correlation matrix/ matrix of scatter plots

### Random Effects

### Reliability

<!-- These functions provide information about the reliability of group-level estimates (i.e., random effects) in mixed models. They are useful to assess whether the predictors yield consistent group-level variability. "Group-level" can refer, for instance, to different participants in a study, and the predictors to the effect of some experimental condition. -->

<!-- D_vour: corresponds to the normalized ratio of observed variability to uncertainty in random effect estimates -->

```{r}
table_dvour <- function(g) {
  performance::performance_dvour(g) |> 
    arrange(desc(D_vour)) |> 
    gt::gt() |> 
    gt::data_color("D_vour", palette = c("red","green"), domain = c(0, 1)) |> 
    gt::fmt_auto()
}
```

```{r}
draft_model <- brms::brm(
  formula = Arousal ~ Gender / (Condition * ConditionBelief) + (1  + Condition | Participant) + (1|Item),
  data = df2,
  algorithm = "meanfield",
  family = brms::zero_one_inflated_beta())

g1 <- modelbased::estimate_grouplevel(draft_model)

# saveRDS(g1, file = "draft_model_estimates.rds")

```

```{r}
# g1 <- readRDS("draft_model_estimates.rds")

table_dvour(g1)

```

#### Participant effects

<!-- The random intercepts for participants showed moderate variability around the grand mean, suggesting that some participants tended to report consistently higher or lower arousal regardless of condition or belief. -->

<!-- The item-level intercepts exhibited similar variability, indicating that certain stimuli were intrinsically more arousing than others across participants. -->

##### Study 1

```{r}
colors <- c("Intercept" = "#9C27B0", "ConditionAIMGenerated" = "#FF9800")

p2_dvour <- g1 |> 
   mutate(Group = fct_relevel(Group, "Participant", "Item"),
          Parameter = fct_relevel(Parameter, "Intercept", "ConditionAIMGenerated")) |>
  ggplot(aes(x = Median , y = Level)) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high, color = Parameter), linewidth = 0.3, size = 0.05) +
  facet_grid(Group~Parameter, scales = "free", switch="both") +
  ggside::ggside(scales  = "free") +
  ggside::geom_xsidedensity(aes(group = Parameter, fill = Parameter), color = NA) +
  ggside::ggside(scales  = "free") +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement.x = "outside",
        strip.background = element_rect(fill = NA, color = NA),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
p2_dvour
```

##### Study 2

```{r}
colors <- c("Intercept" = "#9C27B0", "ConditionAIMGenerated" = "#FF9800")

p2_dvour <- g1 |> 
   mutate(Group = fct_relevel(Group, "Participant", "Item"),
          Parameter = fct_relevel(Parameter, "Intercept", "ConditionAIMGenerated")) |>
  ggplot(aes(x = Median , y = Level)) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high, color = Parameter), linewidth = 0.3, size = 0.05) +
  facet_grid(Group~Parameter, scales = "free", switch="both") +
  ggside::ggside(scales  = "free") +
  ggside::geom_xsidedensity(aes(group = Parameter, fill = Parameter), color = NA) +
  ggside::ggside(scales  = "free") +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement.x = "outside",
        strip.background = element_rect(fill = NA, color = NA),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
p2_dvour
```

### Correlation

- between random-slopes (effect of condition) and dispositional traits 
<!-- - look at mint and fiction choco analysis for correlation matrix -->

#### Study 1

```{r}

df_phase <- as.data.frame(g1) |>
  filter(Group == "Participant") |>
  mutate(Parameter = as.factor(Parameter)) |>
   pivot_wider(id_cols = Level, names_from = Parameter, values_from = c(Median, MAD, CI_low, CI_high),names_glue = "{Parameter}_{.value}"
  ) |>
  rename(Participant= Level)

head(g2_phase)

df_dispositional <- dfsub1 |>
  select(Participant, 
           BAIT_Videos, BAIT_Visual, BAIT_Text, AI_Knowledge, SexualActivity, 
           GAAIS_Negative, GAAIS_Positive, Porn) |>
  rename(Porn_factor = Porn)

head(g3_dispositional)

```

```{r}
make_cor <- function(df_phase, df_dispositional = NULL, 
                     qname_x = TRUE, qname_y = TRUE, 
                     textsize = 3, xcol = "#424242", ycol = "#424242") {

  # Identify common participants (if df_dispositional provided)
  if (!is.null(df_dispositional)) {
    common_ids <- intersect(df_phase$Participant, df_dispositional$Participant)
    df_phase <- df_phase |> filter(Participant %in% common_ids) |> arrange(Participant)
    df_dispositional <- df_dispositional |> filter(Participant %in% common_ids) |> arrange(Participant)
  }

  # Compute correlations
  # r <- df_phase |>
  #   select(ConditionAIMGenerated_Median, Intercept_Median) |>
  #   correlation::correlation(
  #     data2 = if (!is.null(df_dispositional)) df_dispositional |> select(-Participant) else NULL,
  #     p_adjust = "none",
  #     redundant = ifelse(is.null(df_dispositional), TRUE, FALSE)
  #   ) |>
  #   correlation::cor_sort() |>
  #   mutate(
  #     label = ifelse(p < .05, paste0(insight::format_value(r, lead_zero = FALSE, zap_small = TRUE)), ""),
  #     Parameter1 = as.factor(Parameter1),
  #     Parameter2 = as.factor(Parameter2)
  #   )

  r <- df_phase |>
  select(ConditionAIMGenerated_Median, Intercept_Median) |>
  correlation::correlation(
    data2 = if (!is.null(df_dispositional)) df_dispositional |> select(-Participant) else NULL,
    p_adjust = "none",
    redundant = ifelse(is.null(df_dispositional), TRUE, FALSE)
  ) |>
  correlation::cor_sort() |>
  mutate(
    # always show r, add '*' if p < .05
    label = paste0(
      insight::format_value(r, lead_zero = FALSE, zap_small = TRUE),
      ifelse(p < .05, "*", "")
    ),
    Parameter1 = as.factor(Parameter1),
    Parameter2 = as.factor(Parameter2)
  )

  # Clear diagonal labels
  r[as.character(r$Parameter1) == as.character(r$Parameter2), "label"] <- ""

  # Axis names
  xnames <- levels(r$Parameter1)
  ynames <- levels(r$Parameter2)

  # xcol <- case_when(
  #   str_detect(xnames, "Condition") ~ "#1565C0",
  #   str_detect(xnames, "Intercept") ~ "#6A1B9A",
  #   .default = xcol
  # )
  # ycol <- case_when(
  #   str_detect(ynames, "BAIT") ~ "#00838F",
  #   str_detect(ynames, "GAAIS") ~ "#C62828",
  #   str_detect(ynames, "AI_Knowledge") ~ "#EF6C00",
  #   .default = ycol
  # )

  # Bold emphasis (optional, minimal)
  xbold <- case_when(
    str_detect(xnames, "Condition") ~ "bold",
    str_detect(xnames, "Intercept") ~ "bold",
    .default = "plain"
  )
  ybold <- case_when(
    str_detect(ynames, "BAIT") ~ "bold",
    str_detect(ynames, "GAAIS") ~ "bold",
    str_detect(ynames, "Porn") ~ "bold",
    str_detect(ynames, "AI") ~ "bold",
    .default = "plain"
  )

  # Plot
  r |>
    ggplot(aes(x = Parameter1, y = Parameter2, fill = r)) +
    geom_tile() +
    geom_text(aes(label = label), size = textsize, lineheight = 0.9) +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(color = xcol, face = xbold),
      axis.text.y = element_text(color = ycol, face = ybold),
      legend.position = "none",
      axis.title = element_blank()
    )
}

p_corAI <- make_cor(df_phase, df_dispositional, qname_x = FALSE, qname_y = FALSE, textsize = 3)  +
  labs(title="Correlations between the effect of condition and dispositional traits") +
  theme(axis.title = element_blank())

p_corAI 
```

#### Study 2